### 2023-02-03 products/manuscript/Manuscript.qmd
Part 1 of MADA Project

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

**PART 1**

**DATA SOURCE AND DATA DESCRIPTION**

The data for the project is extracted from the National Health Interview Survey (NHIS). The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S. population and makes digital data files available for years 1963 to present. Users can create their own costumed NHIS data extracts for analysis from the given website: <https://nhis.ipums.org/nhis-action/variables/group>.

List of variables extracted in the dataset to be used for the study:

1.  Demographics: age, sex, race/ethnicity, marital status, veteran status, family

2.  Socio-economic status: education, family income, occupation

3.  General health: Health status, BMI

4.  Women health: pregnancy status, hormone replacement therapy status, use of birth control pills/implants/shots

5.  Cancer testing: mammogram screening

6.  Cancer family history: family members with cancer

7.  Health conditions: cancer, diabetes, high cholesterol

8.  Health behaviors: Alcohol consumption, smoking and pregnancy, physical activity

9.  Health insurance

10. Medical care and access

The extracted data set contains the above listed variables information for years 2000, 2005, 2010, and 2015. As most data were missing for most of the other years, I decided to go with the selected years (2000, 2005, 2010, and 2015) as there were most data points available for those years and 5-years difference comparison might give great insights at the trends of mammogram screening over years. The extracted data set has 393,032 rows and 138 columns.

**QUESTIONS/HYPOTHESIS TO BE ADDRESSED**

The main research question of this study is: Is there any association between mammogram screening for breast cancer and socio-economic status of people?

**METHOD OF ANALYSIS**

Multiple regressions models with robust standard errors will be used for examining the association between mammogram screening and socio-economic status of people adjusting for age, sex, race/ethnicity, family history with cancer, and other control variables. Predictor variables will be various socio-economic determinant variables while the outcome variable will be mammogram screening for breast cancer.

**---------------------------------------------------------------------------------------------------------------------------------**

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

*Provide enough background on your topic that others can understand the why and how of your analysis*

## Description of data and data source

*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2023-02-24 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
Part 2 of MADA project

---
title: "Codes_for_EDA"
author: "Shiwani Sapkota"
date: "2023-02-24"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(gmodels) #for ch-sq freuency table
library(vcd) # for three way frequency table
library(gtsummary)
```

```{r}
# Reading the cleaned csv file
data_clean <- readRDS("mammogram_ses_clean.rds")
head(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
gg_miss_var(data_clean_main)

# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

CREATING DEMOGRAPHIC CHARACTERISTICS TABLE

```{r}
# Lets look at the distribution of hipanic and non-hispanic women during the given years
CrossTable(data_clean_main$year, data_clean_main$hispanic_status)

# Lets look at the mammogram status among hipanic and non-hispanic women during the given years
CrossTable(data_clean_main$mammogram_status, data_clean_main$hispanic_status)

# Lets look at the number of mammogram status of women aged 40-74 during the years
CrossTable(data_clean_main$year, data_clean_main$mammogram_status)

# Lets look only at the socio-economic characteristics of women aged 40-64 by hispanic race status for different years
# Lets look at the age distribution among hispanic and non-hipanic women during the given years
year_age_hisp <- data_clean_main %>% select(year, age, hispanic_status)
mytable1 <- xtabs(~year+age+hispanic_status, data = year_age_hisp)
ftable(mytable1)
ftable(addmargins(prop.table(mytable1, c(1, 2)), 3))*100

# Lets look at the mammogram distribution among hispanic and non-hipanic women during the given years
year_mammogram_hisp <- data_clean_main %>% select(year, mammogram_status, hispanic_status)
mytable2 <- xtabs(~year+mammogram_status+hispanic_status, data = year_mammogram_hisp)
ftable(mytable2)
ftable(addmargins(prop.table(mytable2, c(1, 2)), 3))*100

# Lets look at the income distribution among hispanic and non-hipanic women during the given years
year_income_hisp <- data_clean_main %>% select(year, income, hispanic_status)
mytable3 <- xtabs(~year+income+hispanic_status, data = year_income_hisp)
ftable(mytable3)
ftable(addmargins(prop.table(mytable3, c(1, 2)), 3))*100

# Lets look at the education status among hispanic and non-hipanic women during the given years
year_education_hisp <- data_clean_main %>% select(year, education, hispanic_status)
mytable4 <- xtabs(~year+education+hispanic_status, data = year_education_hisp)
ftable(mytable4)
ftable(addmargins(prop.table(mytable4, c(1, 2)), 3))*100

# Lets look at the marital status among hispanic and non-hipanic women during the given years
year_marital_hisp <- data_clean_main %>% select(year, marital_status, hispanic_status)
mytable5 <- xtabs(~year+marital_status+hispanic_status, data = year_marital_hisp)
ftable(mytable5)
ftable(addmargins(prop.table(mytable5, c(1, 2)), 3))*100

# Lets look at the region of residence among hispanic and non-hipanic women during the given years
year_residence_hisp <- data_clean_main %>% select(year, region_residence, hispanic_status)
mytable6 <- xtabs(~year+region_residence+hispanic_status, data = year_residence_hisp)
ftable(mytable6)
ftable(addmargins(prop.table(mytable6, c(1, 2)), 3))*100

# Lets look at the insurance status among hispanic and non-hipanic women during the given years
year_insurance_hisp <- data_clean_main %>% select(year, insurance_status, hispanic_status)
mytable7 <- xtabs(~year+insurance_status+hispanic_status, data = year_insurance_hisp)
ftable(mytable7)
ftable(addmargins(prop.table(mytable7, c(1, 2)), 3))*100

# Lets look at the usual medical care status among hispanic and non-hipanic women during the given years
year_usualmedicalcare_hisp <- data_clean_main %>% select(year, usual_medicalcare_status, hispanic_status)
mytable8 <- xtabs(~year+usual_medicalcare_status+hispanic_status, data = year_usualmedicalcare_hisp)
ftable(mytable8)
ftable(addmargins(prop.table(mytable8, c(1, 2)), 3))*100

# Lets look at the smoking status among hispanic and non-hipanic women during the given years
year_smoking_hisp <- data_clean_main %>% select(year, smoking_status, hispanic_status)
mytable9 <- xtabs(~year+smoking_status+hispanic_status, data = year_smoking_hisp)
ftable(mytable9)
ftable(addmargins(prop.table(mytable9, c(1, 2)), 3))*100

# Lets look at the alcohol status among hispanic and non-hipanic women during the given years
year_alcohol_hisp <- data_clean_main %>% select(year, alcohol_status, hispanic_status)
mytable10 <- xtabs(~year+alcohol_status+hispanic_status, data = year_alcohol_hisp)
ftable(mytable10)
ftable(addmargins(prop.table(mytable10, c(1, 2)), 3))*100

# Lets look at the diabetes status among hispanic and non-hipanic women during the given years
year_diabetes_hisp <- data_clean_main %>% select(year, diabetes_status, hispanic_status)
mytable11 <- xtabs(~year+diabetes_status+hispanic_status, data = year_diabetes_hisp)
ftable(mytable11)
ftable(addmargins(prop.table(mytable11, c(1, 2)), 3))*100
```

CREATING SOME FIGURES

```{r}

```

### 2023-02-24 code/processing_cleaning_code/Codes_for_cleaning.Rmd
Part 2 of MADA project

---
title: "Codes_for_cleaning"
author: "Shiwani Sapkota"
date: "2023-02-24"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r}
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
```

LOADING DATA FILE

```{r}
# I will be working on rds file so loading and looking at the raw rds data file
setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
data_main <- readRDS("mammogram_ses1.rds")
```

LOOKING AT THE DATA

```{r}
# Looking at the data
glimpse(data_main)
summary(data_main)
skim(data_main)
```

DATA CLEANING

```{r}
# Converting the data file to data frame and converting all columns to lowercase
data_main_df <- data.frame(data_main)
names(data_main_df) <- tolower(names(data_main_df))

# Lets look at the columns to check if our above code worked
colnames(data_main_df)

# Restricting our study to only females i.e. 2
data_main_df <- data_main_df[(which(data_main_df$sex == 2)), ]

# Checking if our above code worked and if there are any missing values in `sex` variable
unique(data_main_df$sex)
dim(data_main_df)
sum(is.na(data_main_df$sex))

# Restricting our study to only females who are 40-74 years old
unique(data_main_df$age)
data_main_df <- subset(data_main_df, data_main_df$age >= 40 & data_main_df$age <= 74)

# Checking if our above code worked and if there are any missing values in `age` variable
unique(data_main_df$age)
dim(data_main_df)
sum(is.na(data_main_df$age))

# Categorizing the `age` variable into 7 categories each with 5 years gap
head(data_main_df$age, n = 5)
data_main_df$age[data_main_df$age >= 40 & data_main_df$age <= 44] <- 1
data_main_df$age[data_main_df$age >= 45 & data_main_df$age <= 49] <- 2
data_main_df$age[data_main_df$age >= 50 & data_main_df$age <= 54] <- 3
data_main_df$age[data_main_df$age >= 55 & data_main_df$age <= 59] <- 4
data_main_df$age[data_main_df$age >= 60 & data_main_df$age <= 64] <- 5
data_main_df$age[data_main_df$age >= 65 & data_main_df$age <= 69] <- 6
data_main_df$age[data_main_df$age >= 70 & data_main_df$age <= 74] <- 7

# Checking to see if our `age` variable was categorized correctly as we wanted
head(data_main_df$age, n = 5)
unique(data_main_df$age)

# Looking at `hispyn` variable (hispanic ethnicity, dichotomous: 1 as non-hispanic and 2 as hipanic ethnicity)
unique(data_main_df$hispyn)
data_main_df <- data_main_df %>% 
  rename("hispanic_status" = "hispyn")
sum(is.na(data_main_df$hispanic_status))
dim(data_main_df)

# Looking at the `pooryn` above or below poverty threshold variable
unique(data_main_df$pooryn)

# Renaming `pooryn` variable to `income`
data_main_df <- data_main_df %>% 
  rename("income" = "pooryn")

# Filtering only at or above poverty threshold: 1 and below poverty threshold:2
data_main_df <- filter(data_main_df, income %in% c(1, 2))
unique(data_main_df$income)
dim(data_main_df)
sum(is.na(data_main_df$income))

# Looking at the `educrec2` education attainment variable
unique(data_main_df$educrec2)

# Renaming `educrec2` variable to `education`
data_main_df <- data_main_df %>% 
  rename("education" = "educrec2")

# Filtering out the `education` variable contents
data_main_df <- filter(data_main_df, education %in% c(10, 20, 31, 32, 41, 42, 51, 54, 60))

# Recoding `education` variable
# education: 1:	Never attended/kindergarten only
#            2:	Less than high school
#            3:	High school
#            4:	Some college or Associate's degree
#            5:	Bachelor's degree and higher
data_main_df$education[data_main_df$education == 10] <- 1
data_main_df$education[data_main_df$education == 20 | data_main_df$education == 31 | data_main_df$education == 32 | data_main_df$education == 41] <- 2
data_main_df$education[data_main_df$education == 42] <- 3
data_main_df$education[data_main_df$education == 51] <- 4
data_main_df$education[data_main_df$education == 54 | data_main_df$education == 60] <- 5
sum(is.na(data_main_df$education))
dim(data_main_df)
unique(data_main_df$education)

# Looking at the `marstcohab` education attainment variable
unique(data_main_df$marstcohab)

# Renaming `marstcohab` variable to `marital_status`
data_main_df <- data_main_df %>% 
  rename("marital_status" = "marstcohab")

# Removing the unknown marital status
data_main_df <- filter(data_main_df, marital_status %in% c(1, 2, 4, 5, 6, 7, 8))
unique(data_main_df$marital_status)

# Recoding `marital_status` variable
# marital_status: 0:	Never married
#                 1:	Married
#                 2:	Single (separated/widowed/divorced)
#                 3:	Living with partner
data_main_df$marital_status[data_main_df$marital_status == 8] <- 0
data_main_df$marital_status[data_main_df$marital_status == 1 | data_main_df$marital_status == 2] <- 1
data_main_df$marital_status[data_main_df$marital_status == 4 | data_main_df$marital_status == 5 | data_main_df$marital_status == 6] <- 2
data_main_df$marital_status[data_main_df$marital_status == 7] <- 3
unique(data_main_df$marital_status)
dim(data_main_df)

# Looking at the `hinotcove` health insurance coverage variable
unique(data_main_df$hinotcove)

# Renaming `hinotcove` variable to `insurance_status`
data_main_df <- data_main_df %>% 
  rename("insurance_status" = "hinotcove")

# Removing the unknown insurance status
data_main_df <- filter(data_main_df, insurance_status %in% c(1, 2))
unique(data_main_df$insurance_status)

# Recoding `insurance_status` variable
# insurance_status: 0:	No coverage
#                   1:	Coverage
data_main_df$insurance_status[data_main_df$insurance_status == 1] <- 0
data_main_df$insurance_status[data_main_df$insurance_status == 2] <- 1
sum(is.na(data_main_df$insurance_status))
unique(data_main_df$insurance_status)
dim(data_main_df)

# Looking at the `hiprivatee` private insurance variable
unique(data_main_df$hiprivatee)

# Renaming `hiprivatee` variable to `privateinsurance_status`
data_main_df <- data_main_df %>% 
  rename("privateinsurance_status" = "hiprivatee")

# Removing the unknown and no information private insurance status
# privateinsurance_status: 1: No
#                          2: Yes
data_main_df <- filter(data_main_df, privateinsurance_status %in% c(1, 2))
sum(is.na(data_main_df$privateinsurance_status))
unique(data_main_df$privateinsurance_status)

# Looking at the `himcaide` medicaid status variable
unique(data_main_df$himcaide)

# Renaming `himcaide` variable to `medicaid_status`
data_main_df <- data_main_df %>% 
  rename("medicaid_status" = "himcaide")

# Removing the unknown and no information medicaid status
# medicaid_status: 1: No
#                  2: Yes
data_main_df <- filter(data_main_df, medicaid_status %in% c(1, 2))
unique(data_main_df$medicaid_status)
sum(is.na(data_main_df$medicaid_status))

# Looking at the `himcaree` medicare status variable
unique(data_main_df$himcaree)

# Renaming `himcaree` variable to `medicare_status`
data_main_df <- data_main_df %>% 
  rename("medicare_status" = "himcaree")

# Removing no information medicare status
# medicare_status: 1: No
#                  2: Yes
data_main_df <- filter(data_main_df, medicare_status %in% c(1, 2))
unique(data_main_df$medicare_status)
sum(is.na(data_main_df$medicare_status))

# Looking at the `health` health status variable
unique(data_main_df$health)

# Renaming `health` variable to `health_status`
data_main_df <- data_main_df %>% 
  rename("health_status" = "health")

# Removing unknown health status and recoding the variable
# health_status: 1: Excellent/very good/good
#                2: Fair or poor
data_main_df <- filter(data_main_df, health_status %in% c(1, 2, 3, 4, 5))
data_main_df$health_status[data_main_df$health_status == 1 | data_main_df$health_status == 2 | data_main_df$health_status == 3] <- 1
data_main_df$health_status[data_main_df$health_status == 4 | data_main_df$marital_status == 5] <- 2
sum(is.na(data_main_df$health_status))
dim(data_main_df)

# Looking at the `mamev` ever had a mammogram variable
unique(data_main_df$mamev)

# Renaming `mamev` variable to `mammogram_status`
data_main_df <- data_main_df %>% 
  rename("mammogram_status" = "mamev")

# Removing unknown mammogram status and recoding the variable
# mammogram_status: 0: No
#                   1: Yes
data_main_df <- filter(data_main_df, mammogram_status %in% c(1, 2))
data_main_df$mammogram_status[data_main_df$mammogram_status == 1] <- 0
data_main_df$mammogram_status[data_main_df$mammogram_status == 2] <- 1
sum(is.na(data_main_df$mammogram_status))
dim(data_main_df)
unique(data_main_df$mammogram_status)

# Looking at the `cancerev` ever told had cancer variable
unique(data_main_df$cancerev)

# Renaming `cancerev` variable to `cancer_status`
data_main_df <- data_main_df %>% 
  rename("cancer_status" = "cancerev")

# Removing unknown cancer status and recoding the variable
# cancer_status: 0: No
#                1: Yes
data_main_df <- filter(data_main_df, cancer_status %in% c(1, 2))
unique(data_main_df$cancer_status)
data_main_df$cancer_status[data_main_df$cancer_status == 1] <- 0
data_main_df$cancer_status[data_main_df$cancer_status == 2] <- 1
sum(is.na(data_main_df$cancer_status))
dim(data_main_df)

# Limiting our study to only non-cancer people
data_main_df <- data_main_df[(which(data_main_df$cancer_status == 0)), ]
unique(data_main_df$cancer_status)
dim(data_main_df)

# Looking at the `smokestatus2` Cigarette smoking recode 2: Current detailed/former/never variable
unique(data_main_df$smokestatus2)

# Renaming `smokestatus2r` variable to `smoking_status`
data_main_df <- data_main_df %>% 
  rename("smoking_status" = "smokestatus2")

# Removing unknown smoking status and recoding the variable
# smoking_status: 0: Never
#                 1: Current
#                 2: Former
data_main_df <- filter(data_main_df, smoking_status %in% c(11, 12, 20, 30))
unique(data_main_df$smoking_status)
data_main_df$smoking_status[data_main_df$smoking_status == 30] <- 0
data_main_df$smoking_status[data_main_df$smoking_status == 11 | data_main_df$smoking_status == 12] <- 1
data_main_df$smoking_status[data_main_df$smoking_status == 20] <- 2
sum(is.na(data_main_df$smoking_status))
dim(data_main_df)

# Looking at the `alcstat1` Alcohol drinking status variable
unique(data_main_df$alcstat1)

# Renaming `alcstat1` variable to `alcohol_status`
data_main_df <- data_main_df %>% 
  rename("alcohol_status" = "alcstat1")

# Removing unknown alcohol status and recoding the variable
# smoking_status: 0: Never
#                 1: Current
#                 2: Former
data_main_df <- filter(data_main_df, alcohol_status %in% c(1:3))
unique(data_main_df$alcohol_status)
data_main_df$alcohol_status[data_main_df$alcohol_status == 1] <- 0
data_main_df$alcohol_status[data_main_df$alcohol_status == 3] <- 1
data_main_df$alcohol_status[data_main_df$alcohol_status == 2] <- 2
sum(is.na(data_main_df$alcohol_status))
dim(data_main_df)

# Looking at the `usualpl` Has usual place for medical care variable
unique(data_main_df$usualpl)

# Renaming `usualpl` variable to `usual_medicalcare_status`
data_main_df <- data_main_df %>% 
  rename("usual_medicalcare_status" = "usualpl")

# Removing unknown usual medicalcare status and recoding the variable
# usual_medicalcare_status: 0: No place
#                           1: Usual place
data_main_df <- filter(data_main_df, usual_medicalcare_status %in% c(1:3))
unique(data_main_df$usual_medicalcare_status)
data_main_df$usual_medicalcare_status[data_main_df$usual_medicalcare_status == 1] <- 0
data_main_df$usual_medicalcare_status[data_main_df$usual_medicalcare_status == 2 | data_main_df$usual_medicalcare_status == 3] <- 1
sum(is.na(data_main_df$usual_medicalcare_status))
dim(data_main_df)

# Looking at the `region` Region of residence variable
unique(data_main_df$region)

# Renaming `region` variable to `region_residence`
# region_residence: 1: Northeast
#                   2: North Central/Midwest
#                   3: South
#                   4: West
data_main_df <- data_main_df %>% 
  rename("region_residence" = "region")
sum(is.na(data_main_df$region_residence))

# Looking at the `diabeticev` Ever told had diabetes variable
unique(data_main_df$diabeticev)

# Renaming `diabeticev` variable to `diabetes_status`
data_main_df <-  data_main_df %>% 
  rename("diabetes_status" = "diabeticev")

# Removing unknown diabetes status and recoding the variable
# diabetes_status: 0: No
#                  1: Yes
data_main_df <- filter(data_main_df, diabetes_status %in% c(1, 2, 3))
data_main_df$diabetes_status[data_main_df$diabetes_status == 1] <- 0
data_main_df$diabetes_status[data_main_df$diabetes_status == 2 | data_main_df$diabetes_status == 3] <- 1
sum(is.na(data_main_df$diabetes_status))
dim(data_main_df)
```

SAVING THE CLEANED DATA FILE AS A CSV and RDS FILE

```{r}
# Commenting this out so that it does not get created many times
# data_main_clean <- subset(data_main_df, select = c("year", "strata", "psu", "perweight", "sampweight", "sex", "age", "hispanic_status", "income", "education", "marital_status", "region_residence", "insurance_status", "privateinsurance_status", "medicaid_status", "medicare_status", "health_status", "mammogram_status", "cancer_status", "usual_medicalcare_status", "smoking_status", "alcohol_status", "diabetes_status"))
# write_csv(data_main_clean, "mammogram_ses_clean.csv")
# saveRDS(data_main_clean, "mammogram_ses_clean.rds")
```

### 2023-02-25 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
Project Part 2 Updates

@@
 library(survey)
-library(gmodels) #for ch-sq freuency table
+library(naniar)
+library(gmodels) #for ch-sq frequency table
 library(vcd) # for three way frequency table
 library(gtsummary)
+library(openxlsx)
+library(ggplot2)
+library(ggthemes)
 ```
+
+LOADING AND LOOKING AT THE DATA
 
@@
 # Reading the cleaned csv file
+setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
 data_clean <- readRDS("mammogram_ses_clean.rds")
@@
 ```{r}
+# Recoding variables to work in making table
+data_clean_maint <- data_clean_main
+data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
+data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
+data_clean_maint$age <- recode_factor(data_clean_maint$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
+data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
+data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
+data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
+data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
+data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
+data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
+data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
+data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'No', '1' = 'Yes')
+
 # Lets look at the distribution of hipanic and non-hispanic women during the given years
-CrossTable(data_clean_main$year, data_clean_main$hispanic_status)
+CrossTable(data_clean_maint$year, data_clean_maint$hispanic_status)
 
 # Lets look at the mammogram status among hipanic and non-hispanic women during the given years
-CrossTable(data_clean_main$mammogram_status, data_clean_main$hispanic_status)
+CrossTable(data_clean_maint$mammogram_status, data_clean_maint$hispanic_status)
 
 # Lets look at the number of mammogram status of women aged 40-74 during the years
-CrossTable(data_clean_main$year, data_clean_main$mammogram_status)
+CrossTable(data_clean_maint$year, data_clean_maint$mammogram_status)
 
@@
 # Lets look at the age distribution among hispanic and non-hipanic women during the given years
-year_age_hisp <- data_clean_main %>% select(year, age, hispanic_status)
+year_age_hisp <- data_clean_maint %>% select(year, age, hispanic_status)
 mytable1 <- xtabs(~year+age+hispanic_status, data = year_age_hisp)
@@
 # Lets look at the mammogram distribution among hispanic and non-hipanic women during the given years
-year_mammogram_hisp <- data_clean_main %>% select(year, mammogram_status, hispanic_status)
+year_mammogram_hisp <- data_clean_maint %>% select(year, mammogram_status, hispanic_status)
 mytable2 <- xtabs(~year+mammogram_status+hispanic_status, data = year_mammogram_hisp)
@@
 # Lets look at the income distribution among hispanic and non-hipanic women during the given years
-year_income_hisp <- data_clean_main %>% select(year, income, hispanic_status)
+year_income_hisp <- data_clean_maint %>% select(year, income, hispanic_status)
 mytable3 <- xtabs(~year+income+hispanic_status, data = year_income_hisp)
@@
 # Lets look at the education status among hispanic and non-hipanic women during the given years
-year_education_hisp <- data_clean_main %>% select(year, education, hispanic_status)
+year_education_hisp <- data_clean_maint %>% select(year, education, hispanic_status)
 mytable4 <- xtabs(~year+education+hispanic_status, data = year_education_hisp)
@@
 # Lets look at the marital status among hispanic and non-hipanic women during the given years
-year_marital_hisp <- data_clean_main %>% select(year, marital_status, hispanic_status)
+year_marital_hisp <- data_clean_maint %>% select(year, marital_status, hispanic_status)
 mytable5 <- xtabs(~year+marital_status+hispanic_status, data = year_marital_hisp)
@@
 # Lets look at the region of residence among hispanic and non-hipanic women during the given years
-year_residence_hisp <- data_clean_main %>% select(year, region_residence, hispanic_status)
+year_residence_hisp <- data_clean_maint %>% select(year, region_residence, hispanic_status)
 mytable6 <- xtabs(~year+region_residence+hispanic_status, data = year_residence_hisp)
@@
 # Lets look at the insurance status among hispanic and non-hipanic women during the given years
-year_insurance_hisp <- data_clean_main %>% select(year, insurance_status, hispanic_status)
+year_insurance_hisp <- data_clean_maint %>% select(year, insurance_status, hispanic_status)
 mytable7 <- xtabs(~year+insurance_status+hispanic_status, data = year_insurance_hisp)
@@
 # Lets look at the usual medical care status among hispanic and non-hipanic women during the given years
-year_usualmedicalcare_hisp <- data_clean_main %>% select(year, usual_medicalcare_status, hispanic_status)
+year_usualmedicalcare_hisp <- data_clean_maint %>% select(year, usual_medicalcare_status, hispanic_status)
 mytable8 <- xtabs(~year+usual_medicalcare_status+hispanic_status, data = year_usualmedicalcare_hisp)
@@
 # Lets look at the smoking status among hispanic and non-hipanic women during the given years
-year_smoking_hisp <- data_clean_main %>% select(year, smoking_status, hispanic_status)
+year_smoking_hisp <- data_clean_maint %>% select(year, smoking_status, hispanic_status)
 mytable9 <- xtabs(~year+smoking_status+hispanic_status, data = year_smoking_hisp)
@@
 # Lets look at the alcohol status among hispanic and non-hipanic women during the given years
-year_alcohol_hisp <- data_clean_main %>% select(year, alcohol_status, hispanic_status)
+year_alcohol_hisp <- data_clean_maint %>% select(year, alcohol_status, hispanic_status)
 mytable10 <- xtabs(~year+alcohol_status+hispanic_status, data = year_alcohol_hisp)
@@
 # Lets look at the diabetes status among hispanic and non-hipanic women during the given years
-year_diabetes_hisp <- data_clean_main %>% select(year, diabetes_status, hispanic_status)
+year_diabetes_hisp <- data_clean_maint %>% select(year, diabetes_status, hispanic_status)
 mytable11 <- xtabs(~year+diabetes_status+hispanic_status, data = year_diabetes_hisp)
@@
 ftable(addmargins(prop.table(mytable11, c(1, 2)), 3))*100
+
+# Lets create a single complete demographic characteristics table by year and hispanic_status of women
+table1 <- data_clean_maint %>% 
+    select(mammogram_status, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, year, hispanic_status) %>%
+    tbl_strata(
+      strata = year,
+      .tbl_fun = 
+        ~ .x %>%
+        tbl_summary(
+        missing = "no",
+        by = c("hispanic_status"),
+        statistic = all_categorical() ~ "{n} ({p}%)"))
+table1
 ```
 
-CREATING SOME FIGURES
+CREATING FIGURE
 
 ```{r}
+# Summarizing mammogram_status and hispanic_status in the given years (Type I)
+data_clean_main1 <- data_clean_main
+data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
 
+# Lets save this figure
+# png("figure1_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
+ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
+  geom_histogram(fill = "white", colour = "red", stat = "count", position = position_dodge()) +
+  facet_grid(hispanic_status ~ year) +
+  scale_x_discrete(labels = c("No", "Yes")) +
+  theme_bw()
+# dev.off()
+
+# Summarizing mammogram_status and hispanic_status in the given years (Type II)
+# png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
+ggplot(data_clean_main1, aes(x = mammogram_status))+
+  geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
+    position = position_dodge(0.9)) +
+  facet_wrap(~year) +
+  scale_x_discrete(labels = c("No", "Yes"))
+# dev.off()
 ```

### 2023-02-25 code/processing_cleaning_code/Codes_for_cleaning.Rmd
Project Part 2 Updates

@@
 data_main_df$health_status[data_main_df$health_status == 1 | data_main_df$health_status == 2 | data_main_df$health_status == 3] <- 1
-data_main_df$health_status[data_main_df$health_status == 4 | data_main_df$marital_status == 5] <- 2
+data_main_df$health_status[data_main_df$health_status == 4 | data_main_df$health_status == 5] <- 2
 sum(is.na(data_main_df$health_status))

### 2023-02-25 products/manuscript/Manuscript.qmd
Project Part 2 Updates

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
-subtitle: ""
-author: Andreas Handel
+title: "Manuscript for MADA Project"
+subtitle: "Mammogram Screening: Sociodemographic factors for hispanic and non-hispanic adult women"
+author: Shiwani Sapkota
 date: "`r Sys.Date()`"
@@
 
-The main research question of this study is: Is there any association between mammogram screening for breast cancer and socio-economic status of people?
+The main research question of this study is: Is there any association between mammogram screening for breast cancer and sociodemographic status of hispanic and non-hispanic adult women?
 

### 2023-02-25 products/supplement/Supplementary_Material.qmd
Project Part 2 Updates

---
title: "Supplement to Manuscript for MADA Project"
subtitle: ""
author: Shiwani Sapkota
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../dataanalysis_template_references.bib
csl: ../../vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(knitr)
library(ipumsr)
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for ch-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
```

# Overview

In this supplement file, you can find details about the data file extraction and codes involved in the process of data extraction and data analysis.

# Code and file information

```{r}
# Skip this code chunk if you already have `mammogram_ses1.rds` raw data file
# NHIS provides .xml and .dat file while extracting the data from website
# Commenting these codes as they were used for data extraction and data preparation process which are needed to be done only once before preparing the data file

# Preparing and loading the raw data file and saving as .csv and .rds files
# setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
# ddi <- read_ipums_ddi("nhis_00003.xml")
# data <- read_ipums_micro(ddi)

# Saving the raw data file as .csv file and commenting out this code so it does not get created everytime we run this code chunk
# readr::write_csv(data, mammogram_ses.csv")

# It seems the created csv file is very large (153 MB) and GitKraken won't like it so lets create an rds file so that could be of small size

# Saving the raw data file as .rds file and commenting out this code so it does not get created everytime we run this code chunk
# saveRDS(data, "mammogram_ses1.rds")

# From here, you can follow general steps involved in data analysis and the
# respective codes are documented in the respected folders of this project:
# 1. Data cleaning
# 2. Exploratory Data analysis
# 3. Main statistical analysis
```

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2023-03-09 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
Update to MADA Project

@@
 library(ggthemes)
+library(here)
 ```
@@
 # Reading the cleaned csv file
-setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
-data_clean <- readRDS("mammogram_ses_clean.rds")
+data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
 head(data_clean)

### 2023-03-09 code/processing_cleaning_code/Codes_for_cleaning.Rmd
Update to MADA Project

@@
 library(naniar)
+library(here)
 ```
@@
 # I will be working on rds file so loading and looking at the raw rds data file
-setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
-data_main <- readRDS("mammogram_ses1.rds")
+data_main <- readRDS(here("data", "raw_data", "mammogram_ses1.rds"))
 ```
@@
 # data_main_clean <- subset(data_main_df, select = c("year", "strata", "psu", "perweight", "sampweight", "sex", "age", "hispanic_status", "income", "education", "marital_status", "region_residence", "insurance_status", "privateinsurance_status", "medicaid_status", "medicare_status", "health_status", "mammogram_status", "cancer_status", "usual_medicalcare_status", "smoking_status", "alcohol_status", "diabetes_status"))
-# write_csv(data_main_clean, "mammogram_ses_clean.csv")
-# saveRDS(data_main_clean, "mammogram_ses_clean.rds")
+# write_csv(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.csv"))
+# saveRDS(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.rds")
 ```

### 2023-03-09 products/supplement/Supplementary_Material.qmd
Update to MADA Project

@@
 # Preparing and loading the raw data file and saving as .csv and .rds files
-# setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
 # ddi <- read_ipums_ddi("nhis_00003.xml")
@@
 # Saving the raw data file as .csv file and commenting out this code so it does not get created everytime we run this code chunk
-# readr::write_csv(data, mammogram_ses.csv")
+# readr::write_csv(data, "mammogram_ses.csv")
 

### 2023-03-17 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
MADA Project Update

@@
 ---
-title: "Codes_for_EDA"
+title: "Codes_for_exploratory_analysis"
 author: "Shiwani Sapkota"
@@
 
-```{r}
+```{r, warning = FALSE, message = FALSE}
 # Loading required packages
@@
 ```{r}
-# Reading the cleaned csv file
+# Reading the cleaned rds file
 data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
@@
 table1
+
+# Lets save the above generated table as a `demographic_table1.rds` table in results folder
+# Commenting this out so that it does not get created many times
+# saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
 ```
@@
 ```{r}
-# Summarizing mammogram_status and hispanic_status in the given years (Type I)
+# Summarizing mammogram_status and Hispanic_status in the given years (Type I)
 data_clean_main1 <- data_clean_main
@@
 
-# Summarizing mammogram_status and hispanic_status in the given years (Type II)
+# Summarizing mammogram_status and Hispanic_status in the given years (Type II)
 # png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)

### 2023-03-17 code/analysis_code/statistical_analysis/Codes_for_statistical_analysis.Rmd
MADA Project Update

---
title: "Codes_for_statistical_analysis"
author: "Shiwani Sapkota"
date: "2023-03-16"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
```

FITTING LOGISTIC REGRESSION

Since our outcome variable `mammogram_status` is of binary type, we will be using bivariate and multivariate logistic regression methods for model fittings.

```{r}
# Setting up the logistic regression
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

BIVARIATE LOGISTIC REGRESSION

```{r}
# Using `mammogram_status` as the outcome and `year` as the predictor
log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
tidy(log_fit_year)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `age` as the predictor
log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
tidy(log_fit_age)
```

```{r}
# Using `mammogram_status` as the outcome and `income` as the predictor
log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
tidy(log_fit_income)
```

```{r}
# Using `mammogram_status` as the outcome and `education` as the predictor
log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)
tidy(log_fit_education)
```

```{r}
# Using `mammogram_status` as the outcome and `marital_status` as the predictor
log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)
tidy(log_fit_maritalstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `region_residence` as the predictor
log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)
tidy(log_fit_regionresidence)
```

```{r}
# Using `mammogram_status` as the outcome and `insurance_status` as the predictor
log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)
tidy(log_fit_insurancestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `health_status` as the predictor
log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)
tidy(log_fit_healthstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `usual_medicalcare_status` as the predictor
log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)
tidy(log_fit_usualmedicalcarestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `smoking_status` as the predictor
log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)
tidy(log_fit_smokingstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `alcohol_status` as the predictor
log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)
tidy(log_fit_alcoholstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `diabetes_status` as the predictor
log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)
tidy(log_fit_diabetesstatus)
```

MULTIVARIATE LOGISTIC REGRESSION

Fitting full model

```{r}
# Using `mammogram_status` as the outcome and the remaining variables as the predictors
log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
table2 <- tidy(log_fit_fullmodel)
table2
saveRDS(table2, file = (here("results", "final_analysis_result", "full_model.rds")))
```

Fitting model using some predictors

```{r}
# Using `mammogram_status` as the outcome and age & hispanic_status as predictors
log_fit_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_age_hispanic)
```

```{r}
# Using `mammogram_status` as the outcome and demographic characteristics variables as predictors
log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
tidy(log_fit_demographic)
```

```{r}
# Using `mammogram_status` as the outcome and the medical characteristics as the predictors
log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
tidy(log_fit_fullmodel)
```

```{r}
# Using `mammogram_status` as the outcome, and smoking & alcohol characteristics as the predictors
log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
tidy(log_fit_smoking_alcohol)
```

### 2023-03-17 products/manuscript/Manuscript.qmd
MADA Project Update

--- 2023-02-03 products/manuscript/Manuscript.qmd
@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
-subtitle: ""
-author: Andreas Handel
+title: "Manuscript for MADA Project"
+subtitle: "Trend in mammogram screening for breast cancer and socio-economic status of hispanic and non-hispanic women during 2000, 2005, 2010, and 2015"
+author: Shiwani Sapkota
 date: "`r Sys.Date()`"
@@
     highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
-csl: ../apa.csl
 ---
-
-**PART 1**
-
-**DATA SOURCE AND DATA DESCRIPTION**
-
-The data for the project is extracted from the National Health Interview Survey (NHIS).
-The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S.
-population and makes digital data files available for years 1963 to present.
-Users can create their own costumed NHIS data extracts for analysis from the given website: <https://nhis.ipums.org/nhis-action/variables/group>.
-
-List of variables extracted in the dataset to be used for the study:
-
-1.  Demographics: age, sex, race/ethnicity, marital status, veteran status, family
-
-2.  Socio-economic status: education, family income, occupation
-
-3.  General health: Health status, BMI
-
-4.  Women health: pregnancy status, hormone replacement therapy status, use of birth control pills/implants/shots
-
-5.  Cancer testing: mammogram screening
-
-6.  Cancer family history: family members with cancer
-
-7.  Health conditions: cancer, diabetes, high cholesterol
-
-8.  Health behaviors: Alcohol consumption, smoking and pregnancy, physical activity
-
-9.  Health insurance
-
-10. Medical care and access
-
-The extracted data set contains the above listed variables information for years 2000, 2005, 2010, and 2015.
-As most data were missing for most of the other years, I decided to go with the selected years (2000, 2005, 2010, and 2015) as there were most data points available for those years and 5-years difference comparison might give great insights at the trends of mammogram screening over years.
-The extracted data set has 393,032 rows and 138 columns.
-
-**QUESTIONS/HYPOTHESIS TO BE ADDRESSED**
-
-The main research question of this study is: Is there any association between mammogram screening for breast cancer and socio-economic status of people?
-
-**METHOD OF ANALYSIS**
-
-Multiple regressions models with robust standard errors will be used for examining the association between mammogram screening and socio-economic status of people adjusting for age, sex, race/ethnicity, family history with cancer, and other control variables.
-Predictor variables will be various socio-economic determinant variables while the outcome variable will be mammogram screening for breast cancer.
-
-**---------------------------------------------------------------------------------------------------------------------------------**
-
-The structure below is one possible setup for a data analysis project (including the course project).
-For a manuscript, adjust as needed.
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
 
@@
 
-*Write a summary of your project.*
-
-{{< pagebreak >}}
+\[\*\*To be added\]
 
@@
 
-## General Background Information
+Breast cancer is one of the most common cancers and the second leading cause of cancer death among women in the US.
+Mammography screening is a screening process where an x-ray image of the breast is taken to detect irregularities in breast tissue which could be the sign of cancer.
+Mammography screening can help detect and treat breast cancer at early stages reducing the mortality rate.
+Hence, mammography screening can play a great role in early breast cancer detection and treatment, and in reduction in the number of deaths related to breast cancer.
+The U.S.
+Preventive Services Task Force (USPSTF) recommends a biennial (every other year) mammography screening for women aged 50 to 74 years and advises to determine the need for mammography screening on an individual basis for women aged 40 to 49 years.
 
-*Provide enough background on your topic that others can understand the why and how of your analysis*
+There are well established racial differences in breast cancer incidence and mortality rates.
+The rate of breast cancer is higher among White women compared to Black and Hispanic women.
+But Black and Hispanic women are often diagnosed at later stages of the cancer and have lower survival rates compared to White women9.
+Apart from the racial differences, there are some consistent findings regarding the positive association between breast cancer screening and SES factors such as income and having a usual/regular source of medical care but there are some conflicting results regarding the effects of SES factors such as education, age, and marital status on breast cancer screening.
 
-## Description of data and data source
+Little is known about the predictors of using cancer screening services among minority women, particularly Hispanics.
+Very few studies have studied the influence of socioeconomic status (SES) and sociodemographic factors on utilizing mammography screening services among Hispanic/non-Hispanic groups of women.
+It is important to understand the influence of women's SES and race/ethnicity on the use of available health care services for improving knowledge about service utilization pattern in women population and for designing appropriate interventions accordingly.
+Hence, for developing appropriate population-based interventions for increasing the use of mammography screening, it is essential to first identify and understand the association and interaction between the use of mammography screening and demographic/socio-economic characteristics of women.
 
-*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*
+This study aims to explore the relationship trend among socioeconomic status and other predictive factors influencing the likelihood of using mammography screening services within Hispanic and non-Hispanic women aged 40 to 74 years during 2000, 2005, 2010, and 2015.
 
-## Questions/Hypotheses to be addressed
-
-*State the research questions you plan to answer with this analysis.*
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]
-
-{{< pagebreak >}}
+\[\*\*References to be added\]
 
@@
 
-*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
+The study used the data from the National Health Interview Survey (NHIS).
+The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S.
+population and makes digital data files available for years 1963 to present.
+Users can create their own costumed NHIS data extracts for analysis from the NHIS website.
+For this study, data with the variables of interest for years: 2000, 2005, 2010, and 2015 were directly extracted from the NHIS website.
+Those specific 4 years were selected for the study as there were most data points available for those years and 5-years difference comparison could give great insights at the trends of mammogram screening over years.
+The study was restricted among Hispanic and non-Hispanic women aged 40-74 years as per the mammography screening recommendation by USPSTF and those age groups women fall under high-risk groups for developing breast cancer.
 
-## Data aquisition
+Mammography screening status (yes or no) was defined as outcome variable of interest for the study.
+Similarly, the following variables were assessed as predictors of mammogram screening: (1) age (40-44, 45-49, 50-54, 55-59, 60-64, 65-69, or 70-74), (2) income (at or above poverty threshold, or below poverty threshold), (3) education (never attended/kindergarten only, less than high school, high school, some college or associate's degree, or bachelor's degree and higher), (4) marital status (never married, married, or single (separated/widowed/divorced)), (5) region of residence (Northeast, North Central/Midwest, South, or West), (6) insurance status (no coverage or coverage), (7) self-reported health status (excellent/very good/good or fair/poor), (8) having a usual place of medical care (no place or have usual place), (9) smoking status (never, current, or former), (10) alcohol status (never, current, or former), and (11) diabetic status (yes or no).
 
-*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*
-
-## Data import and cleaning
-
-*Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along.*
-
-## Statistical analysis
-
-*Explain anything related to your statistical analyses.*
-
-{{< pagebreak >}}
+The study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015.
+Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24848 subjects.
+\[\*\*References and Statistical analysis part to be added\]
 
@@
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+**Table 1: Demographic characteristics of study population**
 
-@tbl-summarytable shows a summary of the data.
+```{r, warning = FALSE, message = FALSE}
 
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
+library(here)
+library(knitr)
+
+# Loading the `demographic_table1.rds` table which was created during the data exploratory process
+result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table1.rds"))
+result_table1
+```
+
+Table 1 shows the summary of mammography screening status and demographic characteristics of the study population.
+
+**Figure 1: Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015**
 
 ```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
+#| label: Figure 1
+#| fig-cap: "Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015."
 #| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
+
+# Loading the `figure1_mammogram_ses.png` figure which was created during the data exploratory process
+include_graphics(here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"))
 ```
 
-## Basic statistical analysis
+Figure 1 shows the summary of mammography screening status and the Hispanic and non-Hispanic status of study population in different years (i.e. 2000, 2005, 2010, and 2015).
 
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
+## Statistical analysis
 
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+**Table 2: Fitting full model using mammogram_status as outcome variable and other variables as predictor variables.**
 
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight_stratified.png"))
+```{r, warning = FALSE, message = FALSE}
+# Loading the `full_model.rds` table which was created during the statistical analysis process
+result_table2 <- readRDS(here("results", "final_analysis_result", "full_model.rds"))
+result_table2
 ```
-
-## Full analysis
-
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
-
-{{< pagebreak >}}
 
@@
 
-## Summary and Interpretation
+\[\*\*Summary, interpretation, strengths, and limitations of the study to be added\]
 
-*Summarize what you did, what you found and what it means.*
+# Conclusions
 
-## Strengths and Limitations
-
-*Discuss what you perceive as strengths and limitations of your analysis.*
-
-## Conclusions
-
-*What are the main take-home messages?*
-
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
-{{< pagebreak >}}
+\[\*\*Main take-home messages from the study to be added\]
 
 # References
+
+\[\*\*To be added\]

### 2023-03-20 code/analysis_code/statistical_analysis/Codes_for_statistical_analysis.Rmd
Updates addition

@@
 table2
-saveRDS(table2, file = (here("results", "final_analysis_result", "full_model.rds")))
+saveRDS(table2, file = (here("results", "statistical_analysis_result", "full_model.rds")))
 ```
@@
 log_fit_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
-tidy(log_fit_age_hispanic)
+table3 <- tidy(log_fit_age_hispanic)
+saveRDS(table3, file = (here("results", "statistical_analysis_result", "age_hispanic_model.rds")))
 ```
@@
 log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
-tidy(log_fit_demographic)
+table4 <- tidy(log_fit_demographic)
+saveRDS(table4, file = (here("results", "statistical_analysis_result", "demographic_model.rds")))
 ```
@@
 log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
-tidy(log_fit_fullmodel)
+table5 <- tidy(log_fit_medical)
+saveRDS(table5, file = (here("results", "statistical_analysis_result", "medical_model.rds")))
 ```
@@
 log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
-tidy(log_fit_smoking_alcohol)
+table6 <- tidy(log_fit_smoking_alcohol)
+saveRDS(table6, file = (here("results", "statistical_analysis_result", "smoking_alcohol_model.rds")))
 ```

### 2023-03-20 products/manuscript/Manuscript.qmd
Updates addition

@@
 ```{r, warning = FALSE, message = FALSE}
-
+#| echo: FALSE
 library(here)
@@
 ```{r, warning = FALSE, message = FALSE}
+#| echo: FALSE
+
 # Loading the `full_model.rds` table which was created during the statistical analysis process
-result_table2 <- readRDS(here("results", "final_analysis_result", "full_model.rds"))
-result_table2
+result_table2 <- readRDS(here("results", "statistical_analysis_result", "full_model.rds"))
+print(result_table2, n = Inf, width = Inf)
 ```

### 2023-03-20 products/supplement/Supplementary_Material.qmd
Updates addition

@@
 title: "Supplement to Manuscript for MADA Project"
-subtitle: ""
+subtitle: "Supplement Materials"
 author: Shiwani Sapkota
@@
 
-{{< pagebreak >}}
-
-# Additional Method Details
-
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
-
-{{< pagebreak >}}
-
-# Additional results
-
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
-
-## Example additional result
-
-@tbl-resulttable1 shows an additional table summarizing a model fit.
+## Additional result
 
 ```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
+#| label: Additional Table 1
+#| tbl-cap: "Fit table using mammogram status as the outcome and age & hispanic_status as predictors."
 #| echo: FALSE
-resulttable1 = readRDS(here("results","resulttable1.rds"))
-knitr::kable(resulttable1)
+additionaltable1 = readRDS(here("results", "statistical_analysis_result", "age_hispanic_model.rds"))
+knitr::kable(additionaltable1)
 ```
 
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
+```{r}
+#| label: Additional Table 2
+#| tbl-cap: "Fit table using mammogram status as the outcome and demographic characteristics variables as predictors."
+#| echo: FALSE
+additionaltable2 = readRDS(here("results", "statistical_analysis_result", "demographic_model.rds"))
+knitr::kable(additionaltable2)
+```
 
 ```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
+#| label: Additional Table 3
+#| tbl-cap: "Fit table using mammogram status as the outcome and the medical characteristics as the predictors."
 #| echo: FALSE
-knitr::include_graphics(here("results","height_weight.png"))
+additionaltable1 = readRDS(here("results", "statistical_analysis_result", "medical_model.rds"))
+knitr::kable(additionaltable3)
 ```
 
-{{< pagebreak >}}
+```{r}
+#| label: Additional Table 4
+#| tbl-cap: "Fit table using mammogram status as the outcome and smoking & alcohol characteristics as the predictors."
+#| echo: FALSE
+additionaltable1 = readRDS(here("results", "statistical_analysis_result", "smoking_alcohol_model.rds"))
+knitr::kable(additionaltable4)
+```
 
@@
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-{{< pagebreak >}}
+[**To be added]
 
 # References
+
+[**To be added]

### 2023-04-07 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
Part 4 Addition

--- 2023-03-17 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
@@
 library(naniar)
-library(gmodels) #for ch-sq frequency table
+library(gmodels) #for chi-sq frequency table
 library(vcd) # for three way frequency table
@@
 library(here)
+here()
 ```
@@
 # Lets save the above generated table as a `demographic_table1.rds` table in results folder
-# Commenting this out so that it does not get created many times
-# saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
+saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
 ```
@@
 # Lets save this figure
-# png("figure1_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
-ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
+figure1 <- ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
   geom_histogram(fill = "white", colour = "red", stat = "count", position = position_dodge()) +
@@
   theme_bw()
-# dev.off()
+ggsave(filename = here("results","exploratory_analysis_result", "figure1_mammogram_ses.png"), plot = figure1)
 
 # Summarizing mammogram_status and Hispanic_status in the given years (Type II)
-# png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
-ggplot(data_clean_main1, aes(x = mammogram_status))+
+figure1_optional<- ggplot(data_clean_main1, aes(x = mammogram_status))+
   geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
@@
   scale_x_discrete(labels = c("No", "Yes"))
-# dev.off()
+
+# Saving this figure
+ggsave(filename = here("results","exploratory_analysis_result", "figure1optional_mammogram_ses.png"), plot = figure1_optional)
 ```

### 2023-04-07 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Part 4 Addition

---
title: "Model Fitting"
author: "Shiwani Sapkota"
date: "2023-03-16"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
```

FITTING LOGISTIC REGRESSION

Since our outcome variable `mammogram_status` is of binary type, we will be using bivariate and multivariate logistic regression methods for model fittings.

```{r}
# Setting up the logistic regression
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

BIVARIATE LOGISTIC REGRESSION

```{r}
# Using `mammogram_status` as the outcome and `year` as the predictor
log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
log_fit_year

# Looking at the summary of log_fit_year
tidy(log_fit_year)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
log_fit_hispanicstatus

# Looking at the summary of log_fit_hispanicstatus
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `age` as the predictor
log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
log_fit_age

# Looking at the summary of log_fit_age
tidy(log_fit_age)
```

```{r}
# Using `mammogram_status` as the outcome and `income` as the predictor
log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
log_fit_income

# Looking at the summary of log_fit_income
tidy(log_fit_income)
```

```{r}
# Using `mammogram_status` as the outcome and `education` as the predictor
log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)
log_fit_education

# Looking at the summary of log_fit_education
tidy(log_fit_education)
```

```{r}
# Using `mammogram_status` as the outcome and `marital_status` as the predictor
log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)
log_fit_maritalstatus

# Looking at the summary of log_fit_maritalstatus
tidy(log_fit_maritalstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `region_residence` as the predictor
log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)
log_fit_regionresidence

# Looking at the summary of log_fit_regionresidence
tidy(log_fit_regionresidence)
```

```{r}
# Using `mammogram_status` as the outcome and `insurance_status` as the predictor
log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)
log_fit_insurancestatus

# Looking at the summary of log_fit_insurancestatus
tidy(log_fit_insurancestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `health_status` as the predictor
log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)
log_fit_healthstatus

# Looking at the summary of log_fit_healthstatus
tidy(log_fit_healthstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `usual_medicalcare_status` as the predictor
log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)
log_fit_usualmedicalcarestatus

# Looking at the summary of log_fit_usualmedicalcarestatus
tidy(log_fit_usualmedicalcarestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `smoking_status` as the predictor
log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)
log_fit_smokingstatus

# Looking at the summary of log_fit_smokingstatus
tidy(log_fit_smokingstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `alcohol_status` as the predictor
log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)
log_fit_alcoholstatus

# Looking at the summary of log_fit_alcoholstatus
tidy(log_fit_alcoholstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `diabetes_status` as the predictor
log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)
log_fit_diabetesstatus

# Looking at the summary of log_fit_diabetesstatus
tidy(log_fit_diabetesstatus)
```

MULTIVARIATE LOGISTIC REGRESSION

Fitting full model

```{r}
# Using `mammogram_status` as the outcome and the remaining variables as the predictors
log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
log_fit_fullmodel

# Looking at the summary of log_fit_fullmodel
table2 <- tidy(log_fit_fullmodel)
table2
saveRDS(table2, file = (here("results", "statistical_analysis_result", "model_fitting_result", "full_model.rds")))
```

Fitting model using some predictors

```{r}
# Using `mammogram_status` as the outcome and year, age, & hispanic_status as predictors
log_fit_year_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
log_fit_year_age_hispanic

# Looking at the summary of log_fit_year_age_hispanic
table3 <- tidy(log_fit_year_age_hispanic)
saveRDS(table3, file = (here("results", "statistical_analysis_result", "model_fitting_result", "year_age_hispanic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and demographic characteristics variables as predictors
log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
log_fit_demographic

# Looking at the summary of log_fit_demographic
table4 <- tidy(log_fit_demographic)
saveRDS(table4, file = (here("results", "statistical_analysis_result", "model_fitting_result", "demographic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and the medical characteristics as the predictors
log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
log_fit_medical

# Looking at the summary of log_fit_medical
table5 <- tidy(log_fit_medical)
saveRDS(table5, file = (here("results", "statistical_analysis_result", "model_fitting_result", "medical_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome, and smoking & alcohol characteristics as the predictors
log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
log_fit_smoking_alcohol

# Looking at the summary of log_fit_smoking_alcohol
table6 <- tidy(log_fit_smoking_alcohol)
saveRDS(table6, file = (here("results", "statistical_analysis_result", "model_fitting_result", "smoking_alcohol_model.rds")))
```

MODEL PERFORMANCE COMPARISON

```{r}
# Comparing all the above models on the basis of model performance
model_performance <- compare_performance(log_fit_year, log_fit_hispanicstatus, log_fit_age, log_fit_income, log_fit_education, log_fit_maritalstatus, log_fit_regionresidence, log_fit_insurancestatus, log_fit_healthstatus, log_fit_usualmedicalcarestatus, log_fit_smokingstatus, log_fit_alcoholstatus, log_fit_diabetesstatus, log_fit_fullmodel, log_fit_year_age_hispanic, log_fit_demographic, log_fit_medical, log_fit_smoking_alcohol)
table7 <- model_performance

# Saving model_performance result
saveRDS(table7, file = (here("results", "statistical_analysis_result", "model_fitting_result", "model_performance_comparison.rds")))
```

### 2023-04-07 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
Part 4 Addition

---
title: "Model Evaluation"
author: "Shiwani Sapkota"
date: "2023-04-07"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
library(broom)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))

# Changing class of mammogram_status variable from integer to factor as we can't use mammogram_status (main outcome) as an integer while preparing recipe and training the model
class(data_clean_main$mammogram_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$year <- as.factor(data_clean_main$year)
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

Let's start working on improving our model fits.

MODEL EVALUATION

DATA SPLITTING

```{r}
# Setting seed
set.seed(123)

# Putting 3/4 of the data into the training set
data_split <- initial_split(data_clean_main, prop = 3/4)

# Creating data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)
```

**FULL MODEL: CREATING RECIPE THAT FITS A LOGISTIC MODEL USING  AS OUTCOME OF INTEREST AND ALL OTHER VARIABLES AS PREDICTORS**

```{r}
# Using mammogram status as a categorical outcome of interest and all other variables of interest as predictors
data_clean_main_rec <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data)

# Fitting the logistic model
data_clean_main_mod <- logistic_reg() %>% 
                     set_engine("glm")

# Modelling workflow for pairing model and recipe 
data_clean_main_wflow <- workflow() %>% 
  add_model(data_clean_main_mod) %>% 
  add_recipe(data_clean_main_rec)
data_clean_main_wflow

# Using the resulting predictors for preparing recipe and training the model
data_clean_main_fit <- 
data_clean_main_wflow %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
data_clean_main_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**FULL MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE}
# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
predict(data_clean_main_fit, test_data)

# Using argument() with the model plus test data for saving them together
data_clean_main_aug <- 
  augment(data_clean_main_fit, test_data)

data_clean_main_aug %>%
  select(mammogram_status, .pred_0, .pred_1)

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug %>% 
  roc_auc(truth = mammogram_status, .pred_0)
```

**ALTERNATIVE MODEL: USING MAMMOGRAM STATUS AS OUTCOME OF INTEREST AND HISPANIC STATUS AS MAIN PREDICTOR**

```{r}
# Using mammogram status as a categorical outcome of interest and hispanic status as main predictor
data_clean_main_rec2 <- recipe(mammogram_status ~ hispanic_status, data = train_data)

# Fitting the logistic model
data_clean_main_mod2 <- logistic_reg() %>% 
                    set_engine("glm")

# Modelling workflow for pairing model and recipe 
data_clean_main_wflow2 <- workflow() %>% 
  add_model(data_clean_main_mod2) %>% 
  add_recipe(data_clean_main_rec2)
data_clean_main_wflow2

# Using the resulting predictors for preparing recipe and training the model
data_clean_main_fit2 <- 
 data_clean_main_wflow2 %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
data_clean_main_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE}
# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
predict(data_clean_main_fit2, test_data)

# Using argument() with the model plus test data for saving them together
data_clean_main_aug2 <- 
  augment(data_clean_main_fit2, test_data)

data_clean_main_aug2 %>%
  select(mammogram_status, .pred_0, .pred_1)

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug2 %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug2 %>% 
  roc_auc(truth = mammogram_status, .pred_0)
```

From the above result, we see that the full model (using mammogram status as the main outcome of interest and other variables as predictors) has ROC-AUC as 0.76 while the alternative model (using mammogram status as the main outcome of interest and Hispanic status as the main predictor) has ROC-AUC as 0.54. So, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.

### 2023-04-07 code/analysis_code/statistical_analysis/3_machine_learning/model_selection.Rmd
Part 4 Addition

--- 2023-04-07 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
@@
 ---
-title: "Model Evaluation"
+title: "Model Selection"
 author: "Shiwani Sapkota"
@@
 
-```{r, warning = FALSE, message = FALSE}
-# Loading required packages
-library(readr)
+```{r, warning = FALSE}
+library(here)
 library(tidyverse)
-library(dplyr)
-library(gtsummary)
+library(tidymodels)
 library(ggplot2)
-library(ggthemes)
-library(here)
-library(tidymodels)
+library(ranger)
+library(rpart)
+library(glmnet)
+library(vip)
+library(rpart.plot)
 library(performance)
-library(vip)
-library(broom)
+library(yardstick)
+library(recipes)
 ```
 
-LOADING AND LOOKING AT THE DATA
+LOADING DATA
 
@@
 
-Let's start working on improving our model fits.
-
-MODEL EVALUATION
-
-DATA SPLITTING
+**DATA SETUP**
 
 ```{r}
-# Setting seed
+# Setting the random seed to 123
 set.seed(123)
 
-# Putting 3/4 of the data into the training set
-data_split <- initial_split(data_clean_main, prop = 3/4)
+# Splitting dataset into 70% training and 30% testing and using mammogram status outcome as stratification
+data_split <- initial_split(data_clean_main, prop = 7/10, strata = mammogram_status)
 
@@
 test_data  <- testing(data_split)
+
+# Creating 5-fold cross-validation, 5 times repeated
+CV_fold_data <- vfold_cv(train_data, v = 5, repeats = 5, strata = mammogram_status)
+
+# Creating a recipe
+recipe_data <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data) %>%
+  step_dummy(all_nominal(), -all_outcomes())
 ```
 
-**FULL MODEL: CREATING RECIPE THAT FITS A LOGISTIC MODEL USING  AS OUTCOME OF INTEREST AND ALL OTHER VARIABLES AS PREDICTORS**
+**NULL MODEL PERFORMANCE**
 
 ```{r}
-# Using mammogram status as a categorical outcome of interest and all other variables of interest as predictors
-data_clean_main_rec <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data)
+# Using train data
+# Creating a recipe for null model
+recipe_null_train <- recipe(mammogram_status ~ 1, data = train_data) %>%
+  step_dummy(all_nominal(), -all_outcomes())
 
-# Fitting the logistic model
-data_clean_main_mod <- logistic_reg() %>% 
-                     set_engine("glm")
+# Creating a logistic model recipe
+recipe_null_logistic <- logistic_reg() %>% 
+  set_engine("glm") %>% 
+  set_mode("classification")
 
-# Modelling workflow for pairing model and recipe 
-data_clean_main_wflow <- workflow() %>% 
-  add_model(data_clean_main_mod) %>% 
-  add_recipe(data_clean_main_rec)
-data_clean_main_wflow
+# Creating workflow pairing model and recipe 
+workflow_null_train <- workflow() %>% 
+  add_model(recipe_null_logistic) %>% 
+  add_recipe(recipe_null_train)
 
-# Using the resulting predictors for preparing recipe and training the model
-data_clean_main_fit <- 
-data_clean_main_wflow %>% 
-  fit(data = train_data)
+# Fitting null model with the folds created from train data 
+train_null <- fit_resamples(workflow_null_train, resamples = CV_fold_data)
 
-# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
-data_clean_main_fit %>% 
-  extract_fit_parsnip() %>% 
-  tidy()
+# Computing RMSE for train data
+metrics_null_train <- collect_metrics(train_null)
+metrics_null_train
 ```
 
-**FULL MODEL: USING TRAINED WORKFLOW TO PREDICT**
-
-```{r, warning = FALSE}
-# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
-predict(data_clean_main_fit, test_data)
-
-# Using argument() with the model plus test data for saving them together
-data_clean_main_aug <- 
-  augment(data_clean_main_fit, test_data)
-
-data_clean_main_aug %>%
-  select(mammogram_status, .pred_0, .pred_1)
-
-# Creating ROC curve and piping to the autoplot() method
-data_clean_main_aug %>% 
-  roc_curve(truth = mammogram_status, .pred_0) %>% 
-  autoplot()
-
-# Using roc_auc() for estimating the area under the curve 
-data_clean_main_aug %>% 
-  roc_auc(truth = mammogram_status, .pred_0)
-```
-
-**ALTERNATIVE MODEL: USING MAMMOGRAM STATUS AS OUTCOME OF INTEREST AND HISPANIC STATUS AS MAIN PREDICTOR**
-
-```{r}
-# Using mammogram status as a categorical outcome of interest and hispanic status as main predictor
-data_clean_main_rec2 <- recipe(mammogram_status ~ hispanic_status, data = train_data)
-
-# Fitting the logistic model
-data_clean_main_mod2 <- logistic_reg() %>% 
-                    set_engine("glm")
-
-# Modelling workflow for pairing model and recipe 
-data_clean_main_wflow2 <- workflow() %>% 
-  add_model(data_clean_main_mod2) %>% 
-  add_recipe(data_clean_main_rec2)
-data_clean_main_wflow2
-
-# Using the resulting predictors for preparing recipe and training the model
-data_clean_main_fit2 <- 
- data_clean_main_wflow2 %>% 
-  fit(data = train_data)
-
-# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
-data_clean_main_fit2 %>% 
-  extract_fit_parsnip() %>% 
-  tidy()
-```
-
-**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**
-
-```{r, warning = FALSE}
-# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
-predict(data_clean_main_fit2, test_data)
-
-# Using argument() with the model plus test data for saving them together
-data_clean_main_aug2 <- 
-  augment(data_clean_main_fit2, test_data)
-
-data_clean_main_aug2 %>%
-  select(mammogram_status, .pred_0, .pred_1)
-
-# Creating ROC curve and piping to the autoplot() method
-data_clean_main_aug2 %>% 
-  roc_curve(truth = mammogram_status, .pred_0) %>% 
-  autoplot()
-
-# Using roc_auc() for estimating the area under the curve 
-data_clean_main_aug2 %>% 
-  roc_auc(truth = mammogram_status, .pred_0)
-```
-
-From the above result, we see that the full model (using mammogram status as the main outcome of interest and other variables as predictors) has ROC-AUC as 0.76 while the alternative model (using mammogram status as the main outcome of interest and Hispanic status as the main predictor) has ROC-AUC as 0.54.
-So, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.
+From the null model, we got ROC AUC as 0.5.

### 2023-04-07 code/processing_cleaning_code/data_cleaning.Rmd
Part 4 Addition

--- 2023-03-09 code/processing_cleaning_code/Codes_for_cleaning.Rmd
@@
 
-```{r}
+```{r, warning = FALSE, message = FALSE}
 library(readr)
@@
 ```{r}
-# Commenting this out so that it does not get created many times
-# data_main_clean <- subset(data_main_df, select = c("year", "strata", "psu", "perweight", "sampweight", "sex", "age", "hispanic_status", "income", "education", "marital_status", "region_residence", "insurance_status", "privateinsurance_status", "medicaid_status", "medicare_status", "health_status", "mammogram_status", "cancer_status", "usual_medicalcare_status", "smoking_status", "alcohol_status", "diabetes_status"))
-# write_csv(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.csv"))
-# saveRDS(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.rds")
+# Saving our cleaned data as csv and rds files
+data_main_clean <- subset(data_main_df, select = c("year", "sex", "age", "hispanic_status", "income", "education", "marital_status", "region_residence", "insurance_status", "privateinsurance_status", "medicaid_status", "medicare_status", "health_status", "mammogram_status", "cancer_status", "usual_medicalcare_status", "smoking_status", "alcohol_status", "diabetes_status"))
+write_csv(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.csv")))
+saveRDS(data_main_clean, file = (here("data", "processed_cleaned_data", "mammogram_ses_clean.rds")))
 ```

### 2023-04-08 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Part 4 Additional Updates

@@
 table2
-saveRDS(table2, file = (here("results", "statistical_analysis_result", "model_fitting_result", "full_model.rds")))
+saveRDS(table2, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds")))
 ```
@@
 table3 <- tidy(log_fit_year_age_hispanic)
-saveRDS(table3, file = (here("results", "statistical_analysis_result", "model_fitting_result", "year_age_hispanic_model.rds")))
+saveRDS(table3, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds")))
 ```
@@
 table4 <- tidy(log_fit_demographic)
-saveRDS(table4, file = (here("results", "statistical_analysis_result", "model_fitting_result", "demographic_model.rds")))
+saveRDS(table4, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds")))
 ```
@@
 table5 <- tidy(log_fit_medical)
-saveRDS(table5, file = (here("results", "statistical_analysis_result", "model_fitting_result", "medical_model.rds")))
+saveRDS(table5, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds")))
 ```
@@
 table6 <- tidy(log_fit_smoking_alcohol)
-saveRDS(table6, file = (here("results", "statistical_analysis_result", "model_fitting_result", "smoking_alcohol_model.rds")))
+saveRDS(table6, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds")))
 ```
@@
 # Saving model_performance result
-saveRDS(table7, file = (here("results", "statistical_analysis_result", "model_fitting_result", "model_performance_comparison.rds")))
+saveRDS(table7, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds")))
 ```

### 2023-04-08 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
Part 4 Additional Updates

@@
   tidy()
+
+# Saving the above coefficients table
+table8 <- data_clean_main_fit
+saveRDS(table8, file = (here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds")))
 ```
@@
   tidy()
+
+# Saving the above coefficients table
+table9 <- data_clean_main_fit2
+saveRDS(table9, file = (here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds")))
 ```

### 2023-04-08 code/analysis_code/statistical_analysis/3_machine_learning/machine_learning.Rmd
Part 4 Additional Updates

---
title: "Model Selection"
author: "Shiwani Sapkota"
date: "2023-04-07"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE}
library(here)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(ranger)
library(rpart)
library(glmnet)
library(vip)
library(rpart.plot)
library(performance)
library(yardstick)
library(recipes)
```

LOADING DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))

# Changing variables as factors
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$year <- as.factor(data_clean_main$year)
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

**DATA SETUP**

```{r}
# Setting the random seed to 123
set.seed(123)

# Splitting dataset into 70% training and 30% testing and using mammogram status outcome as stratification
data_split <- initial_split(data_clean_main, prop = 7/10, strata = mammogram_status)

# Creating data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)

# Creating 5-fold cross-validation, 5 times repeated
CV_fold_data <- vfold_cv(train_data, v = 5, repeats = 5, strata = mammogram_status)

# Creating a recipe
recipe_data <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())
```

**NULL MODEL PERFORMANCE**

```{r}
# Using train data
# Creating a recipe for null model
recipe_null_train <- recipe(mammogram_status ~ 1, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Creating a logistic model recipe
recipe_null_logistic <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

# Creating workflow pairing model and recipe 
workflow_null_train <- workflow() %>% 
  add_model(recipe_null_logistic) %>% 
  add_recipe(recipe_null_train)

# Fitting null model with the folds created from train data 
train_null <- fit_resamples(workflow_null_train, resamples = CV_fold_data)

# Computing ROC AUC for train data
metrics_null_train <- collect_metrics(train_null)
metrics_null_train
```

From the null model, we got ROC AUC as 0.5.

DECISION TREE MODEL: MODEL TUNING AND FITTING

```{r}
# Tuning hyperparameters by creating model specification that identifies which hyperparameters we are planning to tune
tune_tree_model <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()) %>%
  set_engine("rpart") %>% 
  set_mode("classification")
tune_tree_model

# Creating a regular grid of values for using some convenience functions for each hyperparameter
tree_grid <- grid_regular(cost_complexity(), tree_depth(), levels = 5)
tree_grid %>% count(tree_depth)

# Tuning the workflow using model specification and recipe and model
tree_wf <- workflow() %>%
  add_model(tune_tree_model) %>%
  add_recipe(recipe_data)

# Tuning using cross-validation and the tune_grid() function
tree_res <- tree_wf %>% 
  tune_grid(resamples = CV_fold_data,
    grid = tree_grid)
```

DECISION TREE MODEL: MODEL EVALUATION

```{r}
# Plotting and saving the above results
decision_tree <- tree_res %>% autoplot()
decision_tree
ggsave(filename = here("results","statistical_analysis_result", "3_machine_learning_result", "figure2_decision_tree_model.png"), plot = decision_tree)

# Getting the best-fit model
tree_res %>%
  show_best()
best_tree <- tree_res %>%
  select_best(metric = "roc_auc")

# Getting the final workflow
final_wf <- tree_wf %>% 
  finalize_workflow(best_tree)

# Fitting to the training data with the final workflow
final_fit <- final_wf %>%
  fit(train_data)

# Plotting the final fit
rpart.plot(extract_fit_parsnip(final_fit)$fit, roundint = FALSE)

# Predicted outcomes
predicted_fit <- predict(final_fit, train_data)
```

From the decision-tree model, we got ROC AUC (Area under the receiver operating characteristic curve) as 0.7.

LASSO MODEL: MODEL TUNING AND FITTING

```{r}
# Building the model
model_lasso <- logistic_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet")

# We will be using the recipe (recipe_data) that we created above 
# Creating the workflow
wf_lasso <- workflow() %>% 
  add_model(model_lasso) %>% 
  add_recipe(recipe_data)

# Model tuning using grid
lasso_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))
lasso_grid %>% top_n(-5)
lasso_grid %>% top_n(5)

# Tuning the model using tune_grid() function
res_lasso <- wf_lasso %>% tune_grid(resamples = CV_fold_data, grid = lasso_grid,
    control = control_grid(verbose = FALSE, save_pred = TRUE), metrics = NULL)
res_lasso %>% collect_metrics()
```

LASSO MODEL: MODEL EVALUATION

```{r}
# Plotting the above results
res_lasso %>% autoplot()

# Getting the best-fit model
res_lasso %>%
  show_best()
best_lasso <- res_lasso %>%
  select_best(metric = "roc_auc")

# Getting the final workflow
final_lasso_wf <- wf_lasso %>% 
  finalize_workflow(best_lasso)

# Fitting to the training data with the final workflow
final_lasso_fit <- final_lasso_wf %>% fit(train_data)

# Plotting and saving the final fit
plot_lasso <- extract_fit_engine(final_lasso_fit)
lasso_model_figure <- plot(plot_lasso, "lambda")
lasso_model_figure
ggsave(filename = here("results","statistical_analysis_result", "3_machine_learning_result", "figure3_lasso_model.png"), plot = lasso_model_figure)

# Predicted outcomes
predicted_lasso_fit <- predict(final_lasso_fit, train_data)

# Looking at the coefficients and penalty for each variable in Lasso model
summary(best_lasso)
coefficients <- final_lasso_fit %>% extract_fit_parsnip() %>% tidy()
final_lasso_fit %>% extract_fit_parsnip() %>% tidy()

# Saving the above coefficients and penalty data in a table
table10 <- final_lasso_fit
saveRDS(table10, file = (here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds")))
```

From the Lasso model, we got ROC AUC as 0.8.

Based on the above results, lasso model performed the best as it had the highest ROC AUC compared to decision tree model. Hence, lasso model was selected.

FINAL EVALUATION

```{r}
# Fitting the final lasso model on split data
final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
final_lasso_test %>% collect_metrics()
```

From the above final model, we got ROC AUC as 0.8.

### 2023-04-08 products/manuscript/Manuscript.qmd
Part 4 Additional Updates

@@
 # Loading the `full_model.rds` table which was created during the statistical analysis process
-result_table2 <- readRDS(here("results", "statistical_analysis_result", "full_model.rds"))
+result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds"))
 print(result_table2, n = Inf, width = Inf)

### 2023-04-08 products/supplement/Supplementary_Material.qmd
Part 4 Additional Updates

@@
 #| echo: FALSE
-additionaltable1 = readRDS(here("results", "statistical_analysis_result", "medical_model.rds"))
+additionaltable3 = readRDS(here("results", "statistical_analysis_result", "medical_model.rds"))
 knitr::kable(additionaltable3)
@@
 #| echo: FALSE
-additionaltable1 = readRDS(here("results", "statistical_analysis_result", "smoking_alcohol_model.rds"))
+additionaltable4 = readRDS(here("results", "statistical_analysis_result", "smoking_alcohol_model.rds"))
 knitr::kable(additionaltable4)
+```
+
+```{r}
+#| label: Additional Table 5
+#| tbl-cap: "Model fitting table using mammogram status as the outcome and demographic characteristics as the predictors."
+#| echo: FALSE
+additionaltable5 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds"))
+knitr::kable(additionaltable5)
+```
+
+```{r}
+#| label: Additional Table 6
+#| tbl-cap: "Model fitting table using mammogram status as the outcome and medical characteristics as the predictors."
+#| echo: FALSE
+additionaltable6 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds"))
+knitr::kable(additionaltable6)
+```
+
+```{r}
+#| label: Additional Table 7
+#| tbl-cap: "Model fitting table using mammogram status as the outcome and smoking and alcohol characteristics as the predictors."
+#| echo: FALSE
+additionaltable7 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds"))
+knitr::kable(additionaltable7)
+```
+
+```{r}
+#| label: Additional Table 8
+#| tbl-cap: "Model fitting table using mammogram status as the outcome and year, age, and hispanic characteristics as the predictors."
+#| echo: FALSE
+additionaltable8 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds"))
+knitr::kable(additionaltable8)
+```
+
+```{r}
+#| label: Additional Table 9
+#| tbl-cap: "Model performance comparison."
+#| echo: FALSE
+additionaltable9 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
+knitr::kable(additionaltable9)
+```
+
+```{r}
+#| label: Additional Table 10
+#| tbl-cap: "Model coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation."
+#| echo: FALSE
+additionaltable10 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
+knitr::kable(additionaltable10)
+```
+
+```{r}
+#| label: Additional Table 11
+#| tbl-cap: "Model coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation."
+#| echo: FALSE
+additionaltable11 = readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
+knitr::kable(additionaltable11)
+```
+
+```{r}
+#| label: Additional Table 12
+#| tbl-cap: "Lasso model final fit using machine learning approach."
+#| echo: FALSE
+additionaltable12 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
+knitr::kable(additionaltable12)
+```
+
+```{r}
+#| label: Additional Figure 1
+#| fig-cap: "Decision tree model."
+#| echo: FALSE
+knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure2_decision_tree_model.png")
+```
+
+```{r}
+#| label: Additional Figure 2
+#| fig-cap: "Lasso model."
+#| echo: FALSE
+knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
 ```

### 2023-04-11 products/manuscript/Manuscript.qmd
Project references addition

@@
 date: "`r Sys.Date()`"
+bibliography: citations.bib
+csl: jama.csl
 format:
@@
 Breast cancer is one of the most common cancers and the second leading cause of cancer death among women in the US.
-Mammography screening is a screening process where an x-ray image of the breast is taken to detect irregularities in breast tissue which could be the sign of cancer.
-Mammography screening can help detect and treat breast cancer at early stages reducing the mortality rate.
+Mammography screening is a screening process where an x-ray image of the breast is taken to detect irregularities in breast tissue which could be the sign of cancer (@gorina_patterns_2021).
+Mammography screening can help detect and treat breast cancer at early stages reducing the mortality rate (@tabar_reduction_1985, @kerlikowske_efficacy_1995, @shapiro_ten-_1982, @laara_trends_1987).
 Hence, mammography screening can play a great role in early breast cancer detection and treatment, and in reduction in the number of deaths related to breast cancer.
 The U.S.
-Preventive Services Task Force (USPSTF) recommends a biennial (every other year) mammography screening for women aged 50 to 74 years and advises to determine the need for mammography screening on an individual basis for women aged 40 to 49 years.
+Preventive Services Task Force (USPSTF) recommends a biennial (every other year) mammography screening for women aged 50 to 74 years and advises to determine the need for mammography screening on an individual basis for women aged 40 to 49 years (@siu_screening_2016).
 
-There are well established racial differences in breast cancer incidence and mortality rates.
+There are well established racial differences in breast cancer incidence and mortality rates (@selvin_breast_2003, @aftab_trends_2021).
 The rate of breast cancer is higher among White women compared to Black and Hispanic women.
-But Black and Hispanic women are often diagnosed at later stages of the cancer and have lower survival rates compared to White women9.
-Apart from the racial differences, there are some consistent findings regarding the positive association between breast cancer screening and SES factors such as income and having a usual/regular source of medical care but there are some conflicting results regarding the effects of SES factors such as education, age, and marital status on breast cancer screening.
+But Black and Hispanic women are often diagnosed at later stages of the cancer and have lower survival rates compared to White women (@selvin_breast_2003).
+Apart from the racial differences, there are some consistent findings regarding the positive association between breast cancer screening and SES factors such as income and having a usual/regular source of medical care but there are some conflicting results regarding the effects of SES factors such as education, age, and marital status on breast cancer screening (@achat_who_2005).
 
 Little is known about the predictors of using cancer screening services among minority women, particularly Hispanics.
-Very few studies have studied the influence of socioeconomic status (SES) and sociodemographic factors on utilizing mammography screening services among Hispanic/non-Hispanic groups of women.
-It is important to understand the influence of women's SES and race/ethnicity on the use of available health care services for improving knowledge about service utilization pattern in women population and for designing appropriate interventions accordingly.
-Hence, for developing appropriate population-based interventions for increasing the use of mammography screening, it is essential to first identify and understand the association and interaction between the use of mammography screening and demographic/socio-economic characteristics of women.
+Very few studies have studied the influence of socioeconomic status (SES) and sociodemographic factors on utilizing mammography screening services among Hispanic/non-Hispanic groups of women (@selvin_breast_2003).
+It is important to understand the influence of women's SES and race/ethnicity on the use of available health care services for improving knowledge about service utilization pattern in women population and for designing appropriate interventions accordingly (@monnat_race/ethnicity_2014).
+Hence, for developing appropriate population-based interventions for increasing the use of mammography screening, it is essential to first identify and understand the association and interaction between the use of mammography screening and demographic/socio-economic characteristics of women (@selvin_breast_2003).
 
 This study aims to explore the relationship trend among socioeconomic status and other predictive factors influencing the likelihood of using mammography screening services within Hispanic and non-Hispanic women aged 40 to 74 years during 2000, 2005, 2010, and 2015.
-
-\[\*\*References to be added\]
 
@@
 The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S.
-population and makes digital data files available for years 1963 to present.
+population and makes digital data files available for years 1963 to present (@blewett_lynn_a._ipums_2016).
 Users can create their own costumed NHIS data extracts for analysis from the NHIS website.
@@
 # References
-
-\[\*\*To be added\]

### 2023-04-21 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
Project Update

--- 2023-03-17 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
@@
 data_clean_maint <- data_clean_main
-data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
+data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'NH', '2' = 'H')
 data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
@@
         by = c("hispanic_status"),
-        statistic = all_categorical() ~ "{n} ({p}%)"))
+        statistic = all_categorical() ~ "{n} ({p}%"))
 table1
 
-# Lets save the above generated table as a `demographic_table1.rds` table in results folder
-# Commenting this out so that it does not get created many times
-# saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
+# Saving the table
+saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
 ```
@@
 ```{r}
-# Summarizing mammogram_status and Hispanic_status in the given years (Type I)
+# Recoding hispanic_status for the figure
 data_clean_main1 <- data_clean_main
@@
 
-# Lets save this figure
-# png("figure1_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
-ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
-  geom_histogram(fill = "white", colour = "red", stat = "count", position = position_dodge()) +
-  facet_grid(hispanic_status ~ year) +
-  scale_x_discrete(labels = c("No", "Yes")) +
-  theme_bw()
-# dev.off()
-
-# Summarizing mammogram_status and Hispanic_status in the given years (Type II)
-# png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
-ggplot(data_clean_main1, aes(x = mammogram_status))+
+# Summarizing mammogram_status and Hispanic_status in the given years
+figure1 <- ggplot(data_clean_main1, aes(x = mammogram_status))+
   geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
@@
   facet_wrap(~year) +
-  scale_x_discrete(labels = c("No", "Yes"))
-# dev.off()
+  scale_x_discrete(labels = c("No", "Yes")) +
+  scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000, 6000))
+figure1
+
+# Saving this figure
+ggsave(filename = here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"), plot = figure1)
 ```

### 2023-04-21 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Project Update

@@
 library(vip)
+library(gt)
 ```
@@
 table7 <- model_performance
+table7
 

### 2023-04-21 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
Project Update

@@
 # Creating ROC curve and piping to the autoplot() method
-data_clean_main_aug %>% 
+roc_fullmodel <- data_clean_main_aug %>% 
   roc_curve(truth = mammogram_status, .pred_0) %>% 
   autoplot()
+roc_fullmodel
+
+# Saving the 'roc_fullmodel' figure
+ggsave(filename = here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure2_roc_fullmodel.png"), plot = roc_fullmodel)
 
@@
   roc_auc(truth = mammogram_status, .pred_0)
+
 ```
@@
 
-**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**
+**ALTERNATIVE MODEL: USING TRAINED WORKFLOW FOR PREDICTION**
 
@@
 # Creating ROC curve and piping to the autoplot() method
-data_clean_main_aug2 %>% 
+roc_altmodel <- data_clean_main_aug2 %>% 
   roc_curve(truth = mammogram_status, .pred_0) %>% 
   autoplot()
+roc_altmodel
+
+# Saving the 'roc_altmodel' figure
+ggsave(filename = here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure3_roc_altmodel.png"), plot = roc_altmodel)
 

### 2023-04-21 code/analysis_code/statistical_analysis/3_machine_learning/machine_learning.Rmd
Project Update

@@
 
+```{r}
+# Let's collect the ROC AUC metric values for different models
+final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
+final_lasso_test %>% collect_metrics()
+```
+
 FINAL EVALUATION
@@
 # Fitting the final lasso model on split data
-final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
-final_lasso_test %>% collect_metrics()
+ml_results <- data.frame(Model = c("Null", "Decision Tree", "Lasso"),
+                      ROC_AUC = c(0.5, 0.7, 0.8))
+ml_results
+
+# Saving the above data frame
+saveRDS(ml_results, file = (here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds")))
 ```

### 2023-04-21 products/manuscript/Manuscript.qmd
Project Update

@@
 
-# Summary/Abstract
+# 1. Summary/Abstract
 
@@
 
-# Introduction
+# 2. Introduction
 
@@
 
-# Methods
+# 3. Methods
 
 The study used the data from the National Health Interview Survey (NHIS).
-The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S.
-population and makes digital data files available for years 1963 to present (@blewett_lynn_a._ipums_2016).
+The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized US population and makes digital data files available for years 1963 to present (@blewett_lynn_a._ipums_2016).
 Users can create their own costumed NHIS data extracts for analysis from the NHIS website.
@@
 
+Data Cleaning:
+
+Analysis:
+
 The study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015.
 Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24848 subjects.
-\[\*\*References and Statistical analysis part to be added\]
 
-# Results
+[** Statistical analysis part to be added]
 
-## Exploratory/Descriptive analysis
+# 4. Results
+
+# 4.1 Exploratory/Descriptive analysis
 
@@
 
-```{r, warning = FALSE, message = FALSE}
+```{r warning = FALSE, message = FALSE}
 #| echo: FALSE
@@
 # Loading the `demographic_table1.rds` table which was created during the data exploratory process
+
 result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table1.rds"))
@@
 
-## Statistical analysis
+# 4.2 Statistical Analysis
 
-**Table 2: Fitting full model using mammogram_status as outcome variable and other variables as predictor variables.**
+# 4.2.1 Model Fitting
+
+**Table 2: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**
 
@@
 
-# Loading the `full_model.rds` table which was created during the statistical analysis process
+# Loading the `full_model.rds` table which was created during the model fitting analysis process
 result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds"))
@@
 ```
+**Table 3: Model Performance Comparisons.**
 
-# Discussion
+```{r, warning = FALSE, message = FALSE}
+#| echo: FALSE
+
+# Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
+result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
+print(result_table2)
+```
+
+# 4.2.2 Model Evaluation
+
+**Figure 2: ROC Curve for the full model**
+
+```{r}
+#| label: Figure 2
+#| fig-cap: "ROC Curve for the full model."
+#| echo: FALSE
+
+# Loading the `figure2_roc_fullmodel.png` figure which was created during the model evaluation process
+include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure2_roc_fullmodel.png"))
+```
+
+**Figure 3: ROC Curve for the alternative model**
+
+```{r}
+#| label: Figure 3
+#| fig-cap: "ROC Curve for the alternative model."
+#| echo: FALSE
+
+# Loading the `figure3_roc_altmodel.png` figure which was created during the model evaluation process
+include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure3_roc_altmodel.png"))
+```
+
+# 4.2.3 Machine Learning
+
+**Table 4: Model Selection using Null, Decision tree and Lasso methods.**
+
+```{r, warning = FALSE, message = FALSE}
+#| echo: FALSE
+
+# Loading the `machinelearning_results.rds` table which was created during the machine learning analysis process
+result_table3 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds"))
+print(result_table3)
+```
+
+# 5 Discussion
 
@@
 
-# Conclusions
+# 6 Conclusions
 
@@
 
-# References
+# 7 References

### 2023-04-21 products/supplement/Supplementary_Material.qmd
Project Update

@@
 #| label: Additional Figure 1
-#| fig-cap: "Decision tree model."
-#| echo: FALSE
-knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure2_decision_tree_model.png")
-```
-
-```{r}
-#| label: Additional Figure 2
 #| fig-cap: "Lasso model."

### 2023-04-22 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
MADA Project Additional Update

@@
 
-# Lets look at the distribution of hipanic and non-hispanic women during the given years
-CrossTable(data_clean_maint$year, data_clean_maint$hispanic_status)
+# Lets look at the summary
+summary(data_clean_maint)
+
+# Lets look at the distribution of hipanic and mammogram status
+CrossTable(data_clean_maint$year, data_clean_maint$mammogram_status)
+
+# Lets look at the hispanic status and mammogram status
+CrossTable(data_clean_maint$hispanic_status, data_clean_maint$mammogram_status)
+
+# Lets look at the income and mammogram status
+CrossTable(data_clean_maint$income, data_clean_maint$mammogram_status)
+
+# Lets look at the education and mammogram status
+CrossTable(data_clean_maint$education, data_clean_maint$mammogram_status)
+
+# Lets look at the marital status and mammogram status
+CrossTable(data_clean_maint$marital_status, data_clean_maint$mammogram_status)
+
+# Lets look at the region of residence and mammogram status
+CrossTable(data_clean_maint$region_residence, data_clean_maint$mammogram_status)
+
+# Lets look at the insurance status and mammogram status
+CrossTable(data_clean_maint$insurance_status, data_clean_maint$mammogram_status)
+
+# Lets look at the health status and mammogram status
+CrossTable(data_clean_maint$health_status, data_clean_maint$mammogram_status)
+
+# Lets look at the usual medical care status and mammogram status
+CrossTable(data_clean_maint$usual_medicalcare_status, data_clean_maint$mammogram_status)
+
+# Lets look at the smoking status and mammogram status
+CrossTable(data_clean_maint$smoking_status, data_clean_maint$mammogram_status)
+
+# Lets look at the alcohol status and mammogram status
+CrossTable(data_clean_maint$alcohol_status, data_clean_maint$mammogram_status)
+
+# Lets look at the diabetes status and mammogram status
+CrossTable(data_clean_maint$diabetes_status, data_clean_maint$mammogram_status)
 
@@
 
+# Creating a demographic characteristics table on the basis of mammogram screening status
+table1 <- data_clean_maint %>% 
+    select(mammogram_status, hispanic_status, year, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status) %>%
+    tbl_strata(
+      strata = mammogram_status,
+      .tbl_fun = 
+        ~ .x %>%
+        tbl_summary(
+           missing = "no"),
+        statistic = all_categorical() ~ "{n} ({p}%)")
+table1
+
 # Lets create a single complete demographic characteristics table by year and hispanic_status of women
-table1 <- data_clean_maint %>% 
+table2 <- data_clean_maint %>% 
     select(mammogram_status, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, year, hispanic_status) %>%
@@
         by = c("hispanic_status"),
-        statistic = all_categorical() ~ "{n} ({p}%"))
-table1
+        statistic = all_categorical() ~ "{n} ({p}%)"))
+table2
 
-# Saving the table
+# Lets create a single complete demographic characteristics table by mammogram_status and hispanic_status of women
+table3 <- data_clean_maint %>% 
+    select(mammogram_status, year, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, hispanic_status) %>%
+    tbl_strata(
+      strata = mammogram_status,
+      .tbl_fun = 
+        ~ .x %>%
+        tbl_summary(
+           missing = "no",
+        by = c("hispanic_status"),
+        statistic = all_categorical() ~ "{n} ({p}%)"))
+table3
+
+# Saving the above 3 tables
 saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
+saveRDS(table2, file = (here("results", "exploratory_analysis_result", "demographic_table2.rds")))
+saveRDS(table3, file = (here("results", "exploratory_analysis_result", "demographic_table3.rds")))
 ```

### 2023-04-22 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
MADA Project Additional Update

@@
 tidy(log_fit_hispanicstatus)
+exp(-0.5624346)
 ```
@@
 table3 <- tidy(log_fit_year_age_hispanic)
+table3
 saveRDS(table3, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds")))
@@
 table4 <- tidy(log_fit_demographic)
+table4
 saveRDS(table4, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds")))
@@
 table5 <- tidy(log_fit_medical)
+table5
 saveRDS(table5, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds")))
@@
 table6 <- tidy(log_fit_smoking_alcohol)
+table6
 saveRDS(table6, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds")))

### 2023-04-22 products/manuscript/Manuscript.qmd
MADA Project Additional Update

@@
 
-\[\*\*To be added\]
+There exists racial differences in mammography screening and breast cancer incidence rates.
+Hispanic and Black women mostly have later stage cancer diagnosis and lower survival rates compared to White women.
+Mammography screening could be of great importance in reducing such later stage breast cancer diagnosis among minority women.
+In that case, it is essential to understand the association between breast cancer screening (mammogram) and socio-economic status of Hispanic and non-Hispanic women.
+This study was conducted to fill up that knowledge gap with the aim of exploring the relationship trend among socioeconomic status and other predictive factors influencing the likelihood of using mammography screening services within Hispanic and non-Hispanic women.
+
+This study used the publicly available self-reported data from the National Health Interview Survey (NHIS).
+The study was limited to 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015.
+After data cleaning, 24,848 subjects were used for subsequent data analysis.
+
+Women aged 40-74 years were reported having mammogram as 84.5%, 85.3%, 87.8%, and 87.5% during years' 2000, 2005, 2010, and 2015 respectively.
+Hispanic women were found having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%).
+All bivariate and multivariate relationships between mammogram status and hypothesized predictors (year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status) were found to be significantly associated with mammogram screening.
+There was a higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
+Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users.
+
+This study findings could be of great importance as health practitioners could identify high risk behavior individuals and recommend appropriate cancer screening services.
+Given the lower overall breast cancer screening rates among Hispanic women, increased education and awareness should be among top priorities for developing appropriate population-based interventions for increasing the use of mammography screening.
+Furthermore, intervention programs should emphasize on educating women and on making uneducated women aware of such cancer screening services.
 
@@
 The study was restricted among Hispanic and non-Hispanic women aged 40-74 years as per the mammography screening recommendation by USPSTF and those age groups women fall under high-risk groups for developing breast cancer.
+Hence, the study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015.
+Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24,848 subjects.
 
@@
 
-Data Cleaning:
-
-Analysis:
-
-The study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015.
-Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24848 subjects.
-
-[** Statistical analysis part to be added]
+The univariate relationship between the mammogram screening and each hypothesized predictors was assessed using logistic regression.
+Multiple logistic regression was also performed for determining the multivariate effect on mammogram screening.
+Model selection process was performed and the final model model was selected based on ROC AUC (Area under the Receiver Operating Characteristic curve) value.
+The model evaluation was performed using 5-fold cross-validation process repeated for 5 times.
+All analysis were performed using R (version 4.2.1) and cross-validation was performed using the tidymodels package.
+The code and data for reproducing results are available on [Github](https://github.com/shiwanisapkota/shiwanisapkota-MADA-project).
 
@@
 
-# 4.1 Exploratory/Descriptive analysis
+# 4.1. Exploratory/Descriptive analysis
+
+Table 1 shows the demographic characteristics of the study population on the basis of mammogram screening status.
+Out of 24,848 subjects, there were 21,466 (86.4%) women doing mammogram screening while there were 3,382 (13.6%) women who did not do mammogram screening.
+Similarly, there were 21,056 (84.7%) non-Hispanic women and 3,792 (15.3%) Hispanic women.
+Larger proportion of women aged 40-44 years did not  go for mammogram screening compared to other aged women.
+Similarly, women with education level of some college or associate degree had higher proportion of mammogram screening.
+Majority of our participants (58.7%) were never smokers.
+Out of 21,056 non-Hispanic women, 18,432 (87.6%) had mammogram screening while out of 3,792 Hispanic women, 3,034 (80%) had mammogram screening.
+Figure 1 shows increasing trend in mammogram screening among women going through 2000 to 2015.
 
 **Table 1: Demographic characteristics of study population**
-
 ```{r warning = FALSE, message = FALSE}
@@
 
-Table 1 shows the summary of mammography screening status and demographic characteristics of the study population.
-
-**Figure 1: Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015**
-
 ```{r}
@@
 ```
+**Figure 1: Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015**
 
-Figure 1 shows the summary of mammography screening status and the Hispanic and non-Hispanic status of study population in different years (i.e. 2000, 2005, 2010, and 2015).
+# 4.2. Statistical Analysis
+# 4.2.1. Model Fitting
 
-# 4.2 Statistical Analysis
+We aimed to explore the bivariate and multivariate relationship between mammogram screening and other hypothesized predictors i.e., year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status.
+Logistic regression was used for examining such relationships.
+In bivariate relationship (Table 2), we found statistically significant relationship between mammogram screening and hispanic status of women and found the log odds of having mammogram screening for Hispanic women as -0.56 compared to non-Hispanic women.
+The log odds of having mammogram was found as 0.25 in year 2015 compared to that of 2000.
+We also found that the log odds of having mammogram increased with age, and education.
+We found the log odds of having mammogram screening for women with income below the poverty threshold as -0.76 compared to women with income at or above the poverty threshold.
+The log odds of having mammogram for married was 0.67 compared to never married women.
+Women having the usual medical place had the log odds of having mammogram of 1.47 compared to those having no usual medical place.
+***add bivariate relation table 2.
+Similarly, Table 3 shows the multivariate relationship between mammogram screening and all other predictors of interests.
+All predictors of interests were found to be significant predictors of mammogram.
 
-# 4.2.1 Model Fitting
+Table 4 shows the model performance comparison which compared the models' performance including various numbers of predictors.
+The best model performance was found for the full model including all predictors of interests (AIC: 16981 and RMSE: 0.32).
+Similarly, based on model performance comparison, second best model was identified as the model including demographic predictors (age, hispanic status, income, education, marital status, and region of residence) that had AIC of 17764.1 and RMSE of 0.33.
 
-**Table 2: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**
-
+**Table 3: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**
 ```{r, warning = FALSE, message = FALSE}
@@
 ```
-**Table 3: Model Performance Comparisons.**
 
+**Table 4: Model Performance Comparisons.**
 ```{r, warning = FALSE, message = FALSE}
@@
 # Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
-result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
-print(result_table2)
+result_table4 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
+print(result_table4)
 ```
 
-# 4.2.2 Model Evaluation
+# 4.2.2. Model Evaluation
 
-**Figure 2: ROC Curve for the full model**
+Model evaluation was performed for the full model including all predictors of interest and the alternative model including only hispanic status of women as the main predictor for mammogram.
+Data splitting was done by putting 3/4 of the data into the training set and recipe was created accordingly for fitting the full and alternative logistic models.
+Figure 2 shows the ROC curve for the full model (using mammogram status as the main outcome of interest and other variables as predictors) having ROC-AUC as 0.76 while Figure 3 shows the ROC curve for the alternative model (using mammogram status as the main outcome of interest and hispanic status as the main predictor) having ROC-AUC as 0.54.
+The larger the area under the curve, better the model is at distinguishing those with the outcomes from those without the outcomes.
+Hence, from the given 2 figures, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.
 
@@
 ```
-
-**Figure 3: ROC Curve for the alternative model**
+**Figure 2: ROC Curve for the full model**
 
@@
 ```
+**Figure 3: ROC Curve for the alternative model**
 
-# 4.2.3 Machine Learning
+# 4.2.3. Machine Learning
 
-**Table 4: Model Selection using Null, Decision tree and Lasso methods.**
+Different modeling techniques were performed for multivariate analysis using Decision-tree method and Lasso method.
+The data was split into 70% training and 30% testing using mammogram status (outcome) as stratification.
+Five-fold cross-validation was also performed which was repeated for 5 times.
+A null model was used for comparing the model performance and for interpreting the predictors' importance.
+The ROC_AUC value for null model was found as 0.5 and this value was used as a basis for comparing Decision tree and Lasso models.
+The Decision tree model had the ROC_AUC value of 0.7 while Lasso model had the ROC_AUC value of 0.8.
+Both models performed better than the null model.
+But out of all the models, Lasso model performed the best and hence, Lasso model was chosen as the best performing model.
 
+**Table 4: Model Selection using Null method, Decision tree method, and Lasso method.**
 ```{r, warning = FALSE, message = FALSE}
@@
 
-# 5 Discussion
+# 5. Discussion
 
-\[\*\*Summary, interpretation, strengths, and limitations of the study to be added\]
+This study aimed to explore the relationship trend among socioeconomic status and other predictive factors influencing the likelihood of using mammography screening services within Hispanic and non-Hispanic women aged 40 to 74 years during 2000, 2005, 2010, and 2015.
+There are very few studies relating to ethnicity and socio-economic characteristics of women for mammogram screening so this study helps to  identify and understand the association and interaction between the use of mammography screening and demographic/socio-economic characteristics of women.
 
-# 6 Conclusions
+In this study, we examined the mammography screening status of women aged 40 to 74 years and their relationship with the hypothesized predictors: year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status.
+During years 2000, 2005, 2010, and 2015, 84.5%, 85.3%, 87.8%, and 87.5% of women aged 40-74 years reported having mammogram.
+Though there was not much increment in mammography rate from 2010 to 2015 but there was some mammography rate increment compared to 2000 and 2005.
+Such increment in mammography rate could be due to people's increased awareness and education, improved access to healthcare, advanced in technology, and changing mammography recommendation guidelines.
+Our findings support previous research as Hispanic women having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%) (@selvin_breast_2003-1).
+All bivariate and multivariate relationships between mammogram status and predictors were found to be significant hence, all hypothesized predictors were found to have strong association with mammogram screening.
+As suggested in previous studies (@katz_breast_2000), we found higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
+Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users.
+Previous studies have also suggested that people engaging in high risk behaviors are less likely to comply with cancer screening guidelines (@holm_health_1999).
+This finding could be of great importance as health practitioners could identify such individuals and recommend appropriate cancer screening services.
+In full model, we found education and age as the predictors having highest significant association with mammogram screening.
+In that case, intervention programs should emphasize on educating women and make uneducated women aware of such cancer screening services.
 
-\[\*\*Main take-home messages from the study to be added\]
+This study has several strengths, including the large national women sample size which contributes to the generalizability of this study's findings.
+The publicly available dataset from NHIS makes the study easily accessible and feasible for this study's reproducibility.
+Furthermore, this study includes women of minority groups especially Hispanic women, who are largely excluded from breast cancer screening studies.
 
-# 7 References
+Nevertheless, there are some limitations of the study.
+There could be some form of bias resulting from the use of self-reported data.
+The study was limited to only 4 years study (2000, 2005, 2010, and 2015) as most of the data points for other years were missing so this study might not have captured the potential confounding variables that might have changed over time.
+
+# 6. Conclusions
+
+This study confirms that there exists some association between mammogram screening and socio-economic status of Hispanic/non-Hispanic women during different years as all the hypothesized predictors were found to be significantly associated with mammogram status.
+We found age, education, income, and smoking/alcohol behavior as the great predictors for having mammogram screening.
+Further research needs to be done to dive more into these hypothesized predictors to see their absolute and relative effects on mammogram screening.
+Furthermore, given the lower overall breast cancer screening rates among Hispanic women, increased education and awareness should be among top priorities for developing appropriate population-based interventions for increasing the use of mammography screening.
+
+# 7. References

### 2023-04-22 products/supplement/Supplementary_Material.qmd
MADA Project Additional Update

@@
 knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
+version
 ```

### 2023-04-24 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
Additional Updates to MADA Project

@@
 data_clean_main1 <- data_clean_main
-data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
+data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic', '2' = 'Hispanic')
 
 # Summarizing mammogram_status and Hispanic_status in the given years
-figure1 <- ggplot(data_clean_main1, aes(x = mammogram_status))+
+figure1 <- ggplot(data_clean_main1, aes(x = mammogram_status)) +
   geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
@@
   scale_x_discrete(labels = c("No", "Yes")) +
-  scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000, 6000))
+  scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000, 6000)) +
+  theme(legend.position = "bottom", 
+        legend.title = element_text(size = 12))
+
 figure1
@@
 # Saving this figure
-ggsave(filename = here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"), plot = figure1)
+ggsave(filename = here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"), plot = figure1, dpi = 300)
 ```

### 2023-04-24 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Additional Updates to MADA Project

@@
 any(is.na(data_clean_main))
+
+# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
+data_clean_main$age <- as.factor(data_clean_main$age)
+data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
+data_clean_main$income <- as.factor(data_clean_main$income)
+data_clean_main$education <- as.factor(data_clean_main$education)
+data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
+data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
+data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
+data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
+data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
+data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
+data_clean_main$health_status <- as.factor(data_clean_main$health_status)
+data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
+data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
+data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
+data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
+data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
+
+# Recoding the variables
+data_clean_main$hispanic_status <- recode_factor(data_clean_main$hispanic_status, '1' = 'Non-Hispanic', '2' = 'Hispanic')
+data_clean_main$mammogram_status <- recode_factor(data_clean_main$mammogram_status, '0' = 'No', '1' = 'Yes')
+data_clean_main$age <- recode_factor(data_clean_main$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
+data_clean_main$income <- recode_factor(data_clean_main$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
+data_clean_main$education <- recode_factor(data_clean_main$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
+data_clean_main$marital_status <- recode_factor(data_clean_main$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
+data_clean_main$region_residence <- recode_factor(data_clean_main$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
+data_clean_main$insurance_status <- recode_factor(data_clean_main$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
+data_clean_main$health_status <- recode_factor(data_clean_main$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
+data_clean_main$usual_medicalcare_status <- recode_factor(data_clean_main$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
+data_clean_main$smoking_status <- recode_factor(data_clean_main$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_main$alcohol_status <- recode_factor(data_clean_main$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_main$diabetes_status <- recode_factor(data_clean_main$diabetes_status, '0' = 'No', '1' = 'Yes')
 ```
@@
 # Looking at the summary of log_fit_year
-tidy(log_fit_year)
+bivariate_year <- tidy(log_fit_year)
+saveRDS(bivariate_year, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_year.rds")))
 ```
@@
 
-# Looking at the summary of log_fit_hispanicstatus
-tidy(log_fit_hispanicstatus)
-exp(-0.5624346)
+# Looking and saving the summary of log_fit_hispanicstatus
+bivariate_hispanic <- tidy(log_fit_hispanicstatus)
+saveRDS(bivariate_hispanic, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_hispanic.rds")))
 ```
@@
 tidy(log_fit_age)
+bivariate_age <- tidy(log_fit_age)
+saveRDS(bivariate_age, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_age.rds")))
 ```
@@
 tidy(log_fit_income)
+bivariate_income <- tidy(log_fit_income)
+saveRDS(bivariate_income, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_income.rds")))
 ```
@@
 tidy(log_fit_education)
+bivariate_education <- tidy(log_fit_education)
+saveRDS(bivariate_education, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_education.rds")))
 ```
@@
 tidy(log_fit_maritalstatus)
+bivariate_maritalstatus <- tidy(log_fit_maritalstatus)
+saveRDS(bivariate_maritalstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_maritalstatus.rds")))
 ```
@@
 tidy(log_fit_regionresidence)
+bivariate_regionresidence <- tidy(log_fit_regionresidence)
+saveRDS(bivariate_regionresidence, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_regionresidence.rds")))
 ```
@@
 tidy(log_fit_insurancestatus)
+bivariate_insurancestatus <- tidy(log_fit_insurancestatus)
+saveRDS(bivariate_insurancestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_insurancestatus.rds")))
 ```
@@
 tidy(log_fit_healthstatus)
+bivariate_healthstatus <- tidy(log_fit_healthstatus)
+saveRDS(bivariate_healthstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_healthstatus.rds")))
 ```
@@
 tidy(log_fit_usualmedicalcarestatus)
+bivariate_usualmedicalcarestatus <- tidy(log_fit_usualmedicalcarestatus)
+saveRDS(bivariate_usualmedicalcarestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_usualmedicalcarestatus.rds")))
 ```
@@
 tidy(log_fit_smokingstatus)
+bivariate_smokingstatus <- tidy(log_fit_smokingstatus)
+saveRDS(bivariate_smokingstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_smokingstatus.rds")))
 ```
@@
 tidy(log_fit_alcoholstatus)
+bivariate_alcoholstatus <- tidy(log_fit_alcoholstatus)
+saveRDS(bivariate_alcoholstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_alcoholstatus.rds")))
 ```
@@
 tidy(log_fit_diabetesstatus)
+bivariate_diabetesstatus <- tidy(log_fit_diabetesstatus)
+saveRDS(bivariate_diabetesstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_diabetesstatus.rds")))
 ```

### 2023-04-24 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
MADA Project Additional Update
Additional Updates to MADA Project

@@
 library(broom)
+library(gridExtra)
+library(cowplot)
 ```
@@
   roc_auc(truth = mammogram_status, .pred_0)
-
 ```
@@
   roc_auc(truth = mammogram_status, .pred_0)
+
+# Combining and saving the above 2 ROC curves for full model and alternative model
+roc_combined <- plot_grid(roc_fullmodel, roc_altmodel, ncol = 2, align = "h", axis = "tb",
+                           labels = c("ROC Curve for the full model", "ROC Curve for the alternative model"),
+                           label_size = 10)
+roc_combined
+ggsave(filename = here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure4_roc_combined.png"), plot = roc_combined)
 ```

### 2023-04-24 code/analysis_code/statistical_analysis/3_machine_learning/machine_learning.Rmd
Additional Updates to MADA Project

@@
 
-# Plotting and saving the final fit
+# Plotting the final fit
 plot_lasso <- extract_fit_engine(final_lasso_fit)
@@
 lasso_model_figure
-ggsave(filename = here("results","statistical_analysis_result", "3_machine_learning_result", "figure3_lasso_model.png"), plot = lasso_model_figure)
 

### 2023-04-24 products/manuscript/Manuscript.qmd
Additional Updates to MADA Project

@@
 csl: jama.csl
+prefer-html: true
 format:
@@
 
-# 1. Summary/Abstract
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# load R packages
+library(here) 
+library(knitr)
+library(ggplot2)
+library(dplyr)
+library(plotly)
+library(gt)
+library(skimr) 
+library(ggpubr)
+library(kableExtra)
+library(gridExtra)
+```
+
+# Summary/Abstract
 
@@
 
-# 2. Introduction
+# Introduction
 
@@
 
-# 3. Methods
+# Methods
 
@@
 
-# 4. Results
+# Results
 
-# 4.1. Exploratory/Descriptive analysis
+## Exploratory/Descriptive analysis
 
-Table 1 shows the demographic characteristics of the study population on the basis of mammogram screening status.
+Table 1 shows the demographic characteristics of the study population on the basis of mammogram screening status (Yes/No).
 Out of 24,848 subjects, there were 21,466 (86.4%) women doing mammogram screening while there were 3,382 (13.6%) women who did not do mammogram screening.
@@
 
-**Table 1: Demographic characteristics of study population**
-```{r warning = FALSE, message = FALSE}
+**Table 1: Demographic characteristics of study population by mammogram screening status**
+```{r, message = FALSE}
 #| echo: FALSE
-library(here)
-library(knitr)
 
 # Loading the `demographic_table1.rds` table which was created during the data exploratory process
-
 result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table1.rds"))
@@
 ```{r}
-#| label: Figure 1
-#| fig-cap: "Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015."
 #| echo: FALSE
@@
 # Loading the `figure1_mammogram_ses.png` figure which was created during the data exploratory process
-include_graphics(here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"))
+knitr::include_graphics(here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"))
 ```
-**Figure 1: Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015**
+**Figure 1: Summary of mammography screening status and the Hispanic status of study population in 2000, 2005, 2010, and 2015.**
 
-# 4.2. Statistical Analysis
-# 4.2.1. Model Fitting
+## Statistical Analysis
+### Model Fitting
 
@@
 Logistic regression was used for examining such relationships.
-In bivariate relationship (Table 2), we found statistically significant relationship between mammogram screening and hispanic status of women and found the log odds of having mammogram screening for Hispanic women as -0.56 compared to non-Hispanic women.
+In bivariate relationship (Supplementary Tables 3-12), we found statistically significant relationship between mammogram screening and hispanic status of women and found the log odds of having mammogram screening for Hispanic women as -0.56 compared to non-Hispanic women.
 The log odds of having mammogram was found as 0.25 in year 2015 compared to that of 2000.
@@
 Women having the usual medical place had the log odds of having mammogram of 1.47 compared to those having no usual medical place.
-***add bivariate relation table 2.
-Similarly, Table 3 shows the multivariate relationship between mammogram screening and all other predictors of interests.
+Similarly, Table 2 shows the multivariate relationship between mammogram screening and all other predictors of interests.
 All predictors of interests were found to be significant predictors of mammogram.
 
-Table 4 shows the model performance comparison which compared the models' performance including various numbers of predictors.
+Table 3 shows the model performance comparison which compared the models' performance including various numbers of predictors.
 The best model performance was found for the full model including all predictors of interests (AIC: 16981 and RMSE: 0.32).
@@
 
-**Table 3: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**
+**Table 2: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**
 ```{r, warning = FALSE, message = FALSE}
@@
 result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds"))
-print(result_table2, n = Inf, width = Inf)
+kable(result_table2 %>% mutate(term = gsub("as\\.factor", "", term)),  align = "c", format.args = list(digits = 3)) %>%
+  row_spec(row = 0, color = "white") %>% kable_styling(full_width = T, bootstrap_options = "striped", font_size = 10)
 ```
 
-**Table 4: Model Performance Comparisons.**
+**Table 3: Model Performance Comparisons.**
 ```{r, warning = FALSE, message = FALSE}
@@
 # Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
-result_table4 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
-print(result_table4)
+result_table3 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
+kable(result_table3 %>% select(Name, AIC, BIC, RMSE),  align = "c", format.args = list(digits = 3)) %>%
+  row_spec(row = 0, color = "white") %>% kable_styling(full_width = T, bootstrap_options = "striped", font_size = 10)
 ```
 
-# 4.2.2. Model Evaluation
+### Model Evaluation
 
@@
 Data splitting was done by putting 3/4 of the data into the training set and recipe was created accordingly for fitting the full and alternative logistic models.
-Figure 2 shows the ROC curve for the full model (using mammogram status as the main outcome of interest and other variables as predictors) having ROC-AUC as 0.76 while Figure 3 shows the ROC curve for the alternative model (using mammogram status as the main outcome of interest and hispanic status as the main predictor) having ROC-AUC as 0.54.
+Figure 2 shows the ROC curve for the full model (using mammogram status as the main outcome of interest and other variables as predictors) having ROC-AUC as 0.76 and the ROC curve for the alternative model (using mammogram status as the main outcome of interest and hispanic status as the main predictor) having ROC-AUC as 0.54.
 The larger the area under the curve, better the model is at distinguishing those with the outcomes from those without the outcomes.
-Hence, from the given 2 figures, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.
+Hence, from the given figure, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.
 
@@
 #| label: Figure 2
-#| fig-cap: "ROC Curve for the full model."
 #| echo: FALSE
 
-# Loading the `figure2_roc_fullmodel.png` figure which was created during the model evaluation process
-include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure2_roc_fullmodel.png"))
+# Loading the `"figure4_roc_combined.png"l.png` figure which was created during the model evaluation process
+include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure4_roc_combined.png"))
 ```
-**Figure 2: ROC Curve for the full model**
 
-```{r}
-#| label: Figure 3
-#| fig-cap: "ROC Curve for the alternative model."
-#| echo: FALSE
+**Figure 3: ROC Curve for the full model and alternative model.**
 
-# Loading the `figure3_roc_altmodel.png` figure which was created during the model evaluation process
-include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure3_roc_altmodel.png"))
-```
-**Figure 3: ROC Curve for the alternative model**
-
-# 4.2.3. Machine Learning
+### Machine Learning
 
@@
 **Table 4: Model Selection using Null method, Decision tree method, and Lasso method.**
-```{r, warning = FALSE, message = FALSE}
+```{r, warning = FALSE, out.width = "100%", message = FALSE}
 #| echo: FALSE
@@
 # Loading the `machinelearning_results.rds` table which was created during the machine learning analysis process
-result_table3 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds"))
-print(result_table3)
+result_table4 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds"))
+kable(result_table4 %>% select(Model, ROC_AUC),  align = "c") %>%
+  row_spec(row = 0, color = "white") %>% kable_styling(full_width = T, bootstrap_options = "striped", font_size = 14, position = "center")
 ```
 
-# 5. Discussion
+# Discussion
 
@@
 
-# 6. Conclusions
+# Conclusions
 
@@
 
-# 7. References
+# References

### 2023-04-24 products/supplement/Supplementary_Material.qmd
Additional Updates to MADA Project

--- 2023-04-21 products/supplement/Supplementary_Material.qmd
@@
     highlight-style: github
-bibliography: ../../dataanalysis_template_references.bib
-csl: ../../vancouver.csl
 ---
@@
 # Saving the raw data file as .csv file and commenting out this code so it does not get created everytime we run this code chunk
-# readr::write_csv(data, "mammogram_ses.csv")
+# readr::write_csv(here("data", "raw_data", "mammogram_ses.csv")
 
@@
 # Saving the raw data file as .rds file and commenting out this code so it does not get created everytime we run this code chunk
-# saveRDS(data, "mammogram_ses1.rds")
+# saveRDS(here("data", "raw_data", "mammogram_ses1.rds")
 
 # From here, you can follow general steps involved in data analysis and the
-# respective codes are documented in the respected folders of this project:
-# 1. Data cleaning
-# 2. Exploratory Data analysis
-# 3. Main statistical analysis
+# respective codes are documented in the `code` folder with the given subfolders:
+# 1. processing_cleaning_code folder
+# 2. analysis_code folder
 ```
 
-## Additional result
+# List of Tables
+# Additional results
+# Exploratory Analysis
 
-```{r}
-#| label: Additional Table 1
-#| tbl-cap: "Fit table using mammogram status as the outcome and age & hispanic_status as predictors."
+# Supplementary Table 1
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable1 = readRDS(here("results", "statistical_analysis_result", "age_hispanic_model.rds"))
-knitr::kable(additionaltable1)
+# Loading the `demographic_table2.rds` table which was created during the data exploratory process
+result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table2.rds"))
+result_table1
 ```
 
-```{r}
-#| label: Additional Table 2
-#| tbl-cap: "Fit table using mammogram status as the outcome and demographic characteristics variables as predictors."
+# Supplementary Table 2
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable2 = readRDS(here("results", "statistical_analysis_result", "demographic_model.rds"))
-knitr::kable(additionaltable2)
+# Loading the `demographic_table3.rds` table which was created during the data exploratory process
+result_table2 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table3.rds"))
+result_table2
 ```
 
-```{r}
-#| label: Additional Table 3
-#| tbl-cap: "Fit table using mammogram status as the outcome and the medical characteristics as the predictors."
+# Model Fitting
+# Bivariate Logistic Regression
+
+# Supplementary Table 3
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable3 = readRDS(here("results", "statistical_analysis_result", "medical_model.rds"))
-knitr::kable(additionaltable3)
+# Loading the `bivariate_mammogram_year.rds` table which was created during the data fitting process
+result_table3 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_year.rds"))
+result_table3
 ```
 
-```{r}
-#| label: Additional Table 4
-#| tbl-cap: "Fit table using mammogram status as the outcome and smoking & alcohol characteristics as the predictors."
+# Supplementary Table 4
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable4 = readRDS(here("results", "statistical_analysis_result", "smoking_alcohol_model.rds"))
-knitr::kable(additionaltable4)
+# Loading the `bivariate_mammogram_hispanic.rds` table which was created during the data fitting process
+result_table4 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_hispanic.rds"))
+result_table4
 ```
 
-```{r}
-#| label: Additional Table 5
-#| tbl-cap: "Model fitting table using mammogram status as the outcome and demographic characteristics as the predictors."
+# Supplementary Table 5
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable5 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds"))
-knitr::kable(additionaltable5)
+# Loading the `bivariate_mammogram_age.rds` table which was created during the data fitting process
+result_table5 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_age.rds"))
+result_table5
 ```
 
-```{r}
-#| label: Additional Table 6
-#| tbl-cap: "Model fitting table using mammogram status as the outcome and medical characteristics as the predictors."
+# Supplementary Table 6
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable6 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds"))
-knitr::kable(additionaltable6)
+# Loading the `bivariate_mammogram_income.rds` table which was created during the data fitting process
+result_table6 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_income.rds"))
+result_table6
 ```
 
-```{r}
-#| label: Additional Table 7
-#| tbl-cap: "Model fitting table using mammogram status as the outcome and smoking and alcohol characteristics as the predictors."
+# Supplementary Table 7
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable7 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds"))
-knitr::kable(additionaltable7)
+# Loading the `bivariate_mammogram_education.rds` table which was created during the data fitting process
+result_table7 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_education.rds"))
+result_table7
 ```
 
-```{r}
-#| label: Additional Table 8
-#| tbl-cap: "Model fitting table using mammogram status as the outcome and year, age, and hispanic characteristics as the predictors."
+# Supplementary Table 8
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable8 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds"))
-knitr::kable(additionaltable8)
+# Loading the `bivariate_mammogram_maritalstatus.rds` table which was created during the data fitting process
+result_table8 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_maritalstatus.rds"))
+result_table8
 ```
 
-```{r}
-#| label: Additional Table 9
-#| tbl-cap: "Model performance comparison."
+# Supplementary Table 9
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable9 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
-knitr::kable(additionaltable9)
+# Loading the `bivariate_mammogram_regionresidence.rds` table which was created during the data fitting process
+result_table9 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_regionresidence.rds"))
+result_table9
 ```
 
-```{r}
-#| label: Additional Table 10
-#| tbl-cap: "Model coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation."
+# Supplementary Table 10
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable10 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
-knitr::kable(additionaltable10)
+# Loading the `bivariate_mammogram_insurancestatus.rds` table which was created during the data fitting process
+result_table10 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_insurancestatus.rds"))
+result_table10
 ```
 
-```{r}
-#| label: Additional Table 11
-#| tbl-cap: "Model coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation."
+# Supplementary Table 11
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable11 = readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
-knitr::kable(additionaltable11)
+# Loading the `bivariate_mammogram_healthstatus.rds` table which was created during the data fitting process
+result_table11 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_healthstatus.rds"))
+result_table11
 ```
 
-```{r}
-#| label: Additional Table 12
-#| tbl-cap: "Lasso model final fit using machine learning approach."
+# Supplementary Table 12
+```{r, message = FALSE}
 #| echo: FALSE
-additionaltable12 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
-knitr::kable(additionaltable12)
+# Loading the `bivariate_mammogram_usualmedicalcarestatus.rds` table which was created during the data fitting process
+result_table12 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_usualmedicalcarestatus.rds"))
+result_table12
 ```
 
+# Supplementary Table 13
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading the `bivariate_mammogram_smokingstatus.rds` table which was created during the data fitting process
+result_table13 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_smokingstatus.rds"))
+result_table13
+```
+
+# Supplementary Table 14
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading the `bivariate_mammogram_alcoholstatus.rds` table which was created during the data fitting process
+result_table14 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_alcoholstatus.rds"))
+result_table14
+```
+
+# Supplementary Table 15
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading the `bivariate_mammogram_diabetesstatus.rds` table which was created during the data fitting process
+result_table15 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_diabetesstatus.rds"))
+result_table15
+```
+
+# Supplementary Table 16
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading fit table using mammogram status as the outcome and year, age, & hispanic_status as predictors
+result_table16 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds"))
+result_table16
+```
+
+# Supplementary Table 17
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading fit table using mammogram status as the outcome and demographic characteristics variables as predictors
+result_table17 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds"))
+result_table17
+```
+
+# Supplementary Table 18
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading fit table using mammogram status as the outcome and the medical characteristics as the predictors
+result_table18 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds"))
+result_table18
+```
+
+# Supplementary Table 19
+```{r, message = FALSE}
+#| echo: FALSE
+# Loading fit table using mammogram status as the outcome and smoking & alcohol characteristics as the predictors
+result_table18 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds"))
+result_table18
+```
+
+# Model Evaluation
+
+# Supplementary Table 20
+```{r, message = FALSE}
+#| echo: FALSE
+ # Loading coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation
+result_table20 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
+result_table20
+```
+
+# Supplementary Table 21
+```{r, message = FALSE}
+#| echo: FALSE
+ # Loading coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation
+result_table21 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
+result_table21
+```
+
+# Machine Learning (Model Selection)
+
+# Supplementary Table 22
+
+```{r, message = FALSE}
+#| echo: FALSE
+ # Loading coefficients and penalty data during lasso model process
+result_table22 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
+result_table22
+```
+
+# List of Figures
+
+# Machine Learning (Model Selection)
+
+# Supplementary Figure 1
 ```{r}
 #| label: Additional Figure 1
-#| fig-cap: "Lasso model."
 #| echo: FALSE
-knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
+
+# Loading figure created using decision tree model for the final fit
+suppl_fig1 <- knitr::include_graphics(here("results", "statistical_analysis_result", "3_machine_learning_result", "figure2_decision_tree_model.png"))
+suppl_fig1
 ```
-
-# Discussion
-
-[**To be added]
-
-# References
-
-[**To be added]

### 2023-04-27 reviews/Part 5 Review - Luisi.qmd
Push part 5 review

---
title: Part5 Project Review 
author: NICOLE LUISI
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Trend in mammogram screening for breast cancer and socio-economic status of hispanic and non-hispanic women during 2000, 2005, 2010, and 2015

Name of project author(s): Shiwani Sapkota

Name of project reviewer: Nicole Luisi

# Specific project content evaluation

## Background, Context and Motivation

### Feedback and Comments

The background is well done and comprehensive. It includes a summary of previous work with references and explains the gap that will be addressed by this analysis. The purpose and goals are clear. The format of the citations is a superscript inside of parentheses, I'm not sure if that was intended.

### Summary assessment

* strong contextualization and motivation

## Question description

### Feedback and Comments

Similar to the background, this is clear and well organized. The description of the NHIS dataset provides evidence that it is well suited to evaluate trends in mammography screening among Hispanic and non-Hispanic women by SES and other factors.

### Summary assessment

* question/hypotheses fully clear

## Data description

### Feedback and Comments

The data source is described very well, including specifics about the variables of interest and the way they were recorded as well as the steps to arrive at the final sample. A codebook does not seem to be included in the files - this did not create any challenges for me as everything was described well and commented, but it may be helpful to provide some of this information in the repository for future use. One small adjustment could be including a direct link to an online codebook. I don't feel strongly about this since a lot of detail was provided in the text.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis

### Feedback and Comments

The data wrangling and exploratory analysis was very thorough. I had no issues reproducing these on my end, and comments were used extensively throughout to explain everything that was being done and why. In the exploratory analysis, some cleanup could be done, particularly in the nicely formatted tables - the current version has a mix of upper and lower case labels on the variables, and some are using original variable names with underscores, etc. Again, this is just a small thing and not very important for the exploratory analysis phase if any final tables for the analysis have corrected this. Figure 1 in this section is very helpful, some extra details such as data labels and descriptions could be added to make this one stand out more as it holds valuable information about the sample.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

This section was also easy to follow. It could be beneficial for this type of write-up to present and report the exponentiated values for odds rather than the estimates/log odds. Similarly, a more simple p-value in the table with fewer decimal places (<0.05) would be easier for a reader to quickly interpret. The logic for the model selection is reasonable. In the write-up a bit more information could be included about the models, maybe just a footnote). The title of the ROC curve on the right is cut off so may need to adjust the scale on that image. It may be helpful to include a summary statement that ties everything together for this section and explains which final model was selected and all the relevant details.

### Summary assessment

* strong and reasonable analysis

## Presentation

### Feedback and Comments

The results are presented well overall, some small details could be adjusted to improve the look and feel. In the main Table 1, the case of the text is not consistent throughout and there are some small errors (no header on education section), use of original variable name (i.e. marital_status), etc. Would do another look over this table to make some small improvements to these details. Labels could also be cleaned up on Figure 1. For the models, the results are presented clearly, for this style of report I would just prefer to have p-values with fewer decimals and perhaps the exponentiated estimates. The fit info is fine, a figure could be considered for that instead of another table, but this is fine as is. It looks like some other interesting work is in the supplemental qmd but there is no HTML file, could consider referencing the supplement in the text to direct people to additional info.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions

### Feedback and Comments

The summary is clear and does a good job bringing everything back together. Strengths and limitations were discussed, and future work proposed.

### Summary assessment

* strong, complete and clear discussion

## Further comments

Overall, really great job on this and an interesting question. The write-up is very thorough, and everything in the code was easy to follow and reproducible. The only thing that could use a little bit more attention for the final version is the main tables and figures, and it's just cleaning up small thing to improve the overall quality of the report, no major issues. Really great work!

# Overall project content evaluation

## Structure

### Feedback and Comments

Everything is structured very well and it was easy to find all of the relevant items. Some additional information could be added to the readme files - I do think that people in this class will be much more in tune with what is going on in this template, and it may be helpful to preapre information as if someone would have no idea. The statistical analysis subfolder does not have any readme, so I do think a bulleted summary of what the code files in that folder are doing would be nice to add.

### Summary assessment

* well structured

## Documentation

### Feedback and Comments

The documentation is thorough, just a few small suggestions for places to add details in the previous sections. Overall, code is very well commented as well as logic for the steps.

### Summary assessment

* fully and well documented

## Reproducibility

### Feedback and Comments

Everything was reproducible and easy to follow. I didn't have any issues running any of the code provided or following the logic/steps taken.

### Summary assessment

* fully reproducible without issues

## Thoroughness

### Feedback and Comments

This was a very thorough analysis with a large dataset. From the exploration to the analysis it is clear that many details were considered. There were no major gaps or issues where I had concerns about the approach or interpretation within the data source.

### Summary assessment

* strong level of thoroughness

## Further comments

No major concerns, I just recommended some minor cleanup of details before the final version. I enjoyed working through it!

### 2023-04-28 reviews/Part 5 Review_Klinker/Project_Review-Klinker.qmd
Review Addition

---
title: Project Review Template 
author: YOUR NAME
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Trend in mammogram screening for breast cancer and socio-economic status of Hispanic and non-Hispanic women during 2000, 2005, 2010, and 2015

Name of project author(s): Shiwani Sapkota

Name of project reviewer: Abbie Klinker

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Great background information, well sourced, and detailed descriptions included about the context of the project itself and the data.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

I'm scoring this section a bit lower, not because the author's description was not clear, but because there's no official "question" or hypothesis outlined in the manuscript. Rather, the author describes the project as exploratory in nature to investigate the relationship trend among socio-demographic factors influencing the likelihood of using mammography screening services.

I don't think this is inadequate, but given how this evaluation section was phrased, I'm not sure if Dr. Handel is looking for something a bit more concrete of a research question/specific hypothesis of the outcomes.

### Summary assessment

* question/hypotheses somewhat explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Noted a bit earlier, but data well described in its source and inclusion criteria.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

WRITE YOUR FEEDBACK HERE

### Summary assessment
* major weaknesses in wrangling and exploratory component
* some weaknesses in wrangling and exploratory component
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

This isn't related to the actual analysis methods used, but I'm so impressed by how you were able to view the decision tree output. I have had so much trouble in accessing metrics and performance of ML models using tidymodels and it looks so cool!

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Table 1 and Graph 1 have a mix of variable names and descriptive names. For Table 2, I'm not sure how necessary it is to list out all factors used versus total model evaluation, especially since no measure of total model evaluation was used, until in comparison with the other single-factor models. Other than that, everything looks great!

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Nicely done!

### Summary assessment

* strong, complete and clear discussion

## Further comments

_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project was really well organized, and I had no trouble finding the files I needed or figuring out the order to run the code!

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

There was a lot of documentation throughout the project, either small notes in the code as comments or next steps outlined through text in the Rmd files.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All code ran with no problem; however, I could not get the actual manuscript to knit, and it seems like a problem with one of the markdown specifications. I'm not 100% sure how to fix it - if I need an extra package on my end or if something's a bit off with the specifications.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Several non-ML and several ML models concered and analyzed.

### Summary assessment

* strong level of thoroughness

## Further comments

Good work! My main feedback is to double check knitting your manuscript with a clear environment and maybe cleaning up variable names in tables, but great project!

### 2023-04-28 part5viewYao/Project_Review_Template.qmd
The only thing I change is to add part 5 review. Some of code not work in my computer. I didn't change them in case some merge conflict. I review your html instead.

# Overview

Title of project:Trend in mammogram screening for breast cancer and socio-economic status of hispanic and non-hispanic women during 2000, 2005, 2010, and 2015

Name of project author(s):
Shiwani Sapkota

Name of project reviewer:
Yao Lu

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given?

### Feedback and Comments
The context of the project is detailed. Comprehensive background is included.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question is fully clear, it's

'This study was conducted to fill up that knowledge gap with the aim of exploring the relationship trend among socioeconomic status and other predictive factors influencing the likelihood of using mammography screening services within Hispanic and non-Hispanic women.'

The question is related to the data.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

There are many data in NHIS. Looks like you didn't put the link in one obvious place. If people miss your data link. New user who are not familiar with your major may need to try hard to find you data. Good thing is you describe it well. So professional people may easy to follow you.

### Summary assessment

* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

data_cleaning.Rmd not work well in my computer. Neither exploratory_analysis.Rmd. In html file, I think there should be some summary way you can show your data. Not a long table. Or maybe you can pick few table which you think important and hide others.

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Some code error in my computer to run the code.

Some of data wrangling is included in model_fitting code. I think you can move them to wrangling part to make your modeling more clear.

Some of data wrangling is included in model_evaluation code. At model_evaluation part. You only tried full model and one 'ALTERNATIVE MODEL' describing few statistical reason. As you tried in logistic regression. There are some large p-value  like 'year2005' with 4.88e- 1, 'year2010'with 1.41e- 4. Why not tried forward or backward model selection to pick a optimal model. Also, if I make to this step. I will try ANOVA analysis, to have look at a F-test. Since there are some categorical variables which including multiple strata. T-test may not be enough. About table 3, Model Performance Comparisons. I think you can try adjusted multivariable logistic regression. Since in real world, we don't have to limit our variables number to 2 or 3. Logistic regression can have an optimal model given the dataset, the alpha threshold and the direction of model selection method.

Some of data wrangling is included in machine_leaning. ROC_AUC result were transfer manually which is a not safe way. I have met some datasets which are tricky. When I come back and double check the data wrangling, manual result transfer make me generate error even I think I am a careful person. Use code to transfer your ROC_AUC result is better.

### Summary assessment

* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Table 1 and figure 1 are good. Table is well looked. But I think ANOVA table is needed. Since like 'year', there are multiple strata. T-test may not enough. Table 3 looks good. But I think backward and forward comparison may be better to put here rather than the current one. If we don't have specific reason why we have to limit the number of variables to 2 or 3. Figure 3 is well looked. But I think there should be several better model than current alternative model.(like backward selection model) Since this alternative model were chosen manually without showing statistic reason.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Discussion about confounder is not mentioned. For example, Alcohol and smoke may be related some other disease. Some people with alcohol or smoke may go to hosipital more often. Then they get test more often.  Self-reported bias were mention ,which is good.

### Summary assessment

* strong, complete and clear discussion

## Further comments

_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._

Very good background description. Modeling you can try using code to transfer your result other than transfer it manually.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

well structured. Though some of code not work in my computer.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

fully and well documented.It will be better if put a link where the data download to a obvious place. I am not sure if there is. But I fail to find that link.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Some of code not work in my computer

### Summary assessment

* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Some code not work in my computer. Wrangling part is good. Optimal multivariable logistic regression is missing. Machine leaning part is good.

### Summary assessment

* strong level of thorougness

## Further comments

_Add any other comments regarding the overall project here. Write anything you think can help your classmate improve their project._

Very good background description and discussion. It will be better if try further modeling. Like comparison between using optimal logistic regression AUC and null model. Or compare backward and forward model selection. I think these can be next steps of your table 3.

### 2023-05-05 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
Final Project Update

@@
 data_clean_maint <- data_clean_main
-data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'NH', '2' = 'H')
+data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'non-hispanic', '2' = 'hispanic')
 data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
 data_clean_maint$age <- recode_factor(data_clean_maint$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
-data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
-data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
-data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
-data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
-data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
-data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
-data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
-data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
-data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
-data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'No', '1' = 'Yes')
+data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'at or above poverty threshold', '2' = 'below poverty threshold')
+data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'never attended/kindergarten only', '2' = 'less than high school', '3' = 'high school', '4' = 'some college or Associate degree', '5' = 'bachelor degree and higher')
+data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'never married', '1' = 'married', '2' = 'single (separated/widowed/divorced)', '3' = 'living with partner')
+data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'northeast', '2' = 'north central/midwest', '3' = 'south', '4' = 'west')
+data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'no coverage', '1' = 'coverage')
+data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'excellent/very good/good', '2' = 'fair/poor')
+data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'no place', '1' = 'usual place')
+data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'never', '1' = 'current', '2' = 'former')
+data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'never', '1' = 'current', '2' = 'former')
+data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'no', '1' = 'yes')
 
@@
 
-# Lets look at the distribution of hipanic and mammogram status
+# Lets look at the distribution of year and mammogram_status
 CrossTable(data_clean_maint$year, data_clean_maint$mammogram_status)
 
-# Lets look at the hispanic status and mammogram status
+# Lets look at the hispanic_status and mammogram_status
 CrossTable(data_clean_maint$hispanic_status, data_clean_maint$mammogram_status)
 
-# Lets look at the income and mammogram status
+# Lets look at the income and mammogram_status
 CrossTable(data_clean_maint$income, data_clean_maint$mammogram_status)
@@
 
-# Lets look at the marital status and mammogram status
+# Lets look at the marital_status and mammogram_status
 CrossTable(data_clean_maint$marital_status, data_clean_maint$mammogram_status)
@@
 
-# Lets look at the insurance status and mammogram status
+# Lets look at the insurance_status and mammogram_status
 CrossTable(data_clean_maint$insurance_status, data_clean_maint$mammogram_status)
 
-# Lets look at the health status and mammogram status
+# Lets look at the health_status and mammogram_status
 CrossTable(data_clean_maint$health_status, data_clean_maint$mammogram_status)
 
-# Lets look at the usual medical care status and mammogram status
+# Lets look at the usual_medicalcare_status and mammogram_status
 CrossTable(data_clean_maint$usual_medicalcare_status, data_clean_maint$mammogram_status)
@@
 
-# Lets look at the alcohol status and mammogram status
+# Lets look at the alcohol_status and mammogram_status
 CrossTable(data_clean_maint$alcohol_status, data_clean_maint$mammogram_status)
 
-# Lets look at the diabetes status and mammogram status
+# Lets look at the diabetes_status and mammogram_status
 CrossTable(data_clean_maint$diabetes_status, data_clean_maint$mammogram_status)
 
-# Lets look at the mammogram status among hipanic and non-hispanic women during the given years
+# Lets look at the mammogram_status among Hispanic and Non-hispanic women (hispanic_status)
 CrossTable(data_clean_maint$mammogram_status, data_clean_maint$hispanic_status)
 
-# Lets look at the number of mammogram status of women aged 40-74 during the years
+# Lets look at the number of mammogram_status of women aged 40-74 during the years
 CrossTable(data_clean_maint$year, data_clean_maint$mammogram_status)
 
-# Lets look only at the socio-economic characteristics of women aged 40-64 by hispanic race status for different years
-# Lets look at the age distribution among hispanic and non-hipanic women during the given years
+# Lets look only at the socio-economic characteristics of women aged 40-64 by hispanic_status for different years
+# Lets look at the age distribution among hispanic_status during the given years
 year_age_hisp <- data_clean_maint %>% select(year, age, hispanic_status)
@@
 
-# Lets look at the mammogram distribution among hispanic and non-hipanic women during the given years
+# Lets look at the mammogram_status distribution among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_mammogram_hisp <- data_clean_maint %>% select(year, mammogram_status, hispanic_status)
@@
 
-# Lets look at the income distribution among hispanic and non-hipanic women during the given years
+# Lets look at the income distribution among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_income_hisp <- data_clean_maint %>% select(year, income, hispanic_status)
@@
 
-# Lets look at the education status among hispanic and non-hipanic women during the given years
+# Lets look at the education among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_education_hisp <- data_clean_maint %>% select(year, education, hispanic_status)
@@
 
-# Lets look at the marital status among hispanic and non-hipanic women during the given years
+# Lets look at the marital_status among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_marital_hisp <- data_clean_maint %>% select(year, marital_status, hispanic_status)
@@
 
-# Lets look at the region of residence among hispanic and non-hipanic women during the given years
+# Lets look at the region_residence among hispanic_status (Hispanic and Non-hispanic women during the given years)
 year_residence_hisp <- data_clean_maint %>% select(year, region_residence, hispanic_status)
@@
 
-# Lets look at the insurance status among hispanic and non-hipanic women during the given years
+# Lets look at the insurance_status among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_insurance_hisp <- data_clean_maint %>% select(year, insurance_status, hispanic_status)
@@
 
-# Lets look at the usual medical care status among hispanic and non-hipanic women during the given years
+# Lets look at the usual_medicalcare_status among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_usualmedicalcare_hisp <- data_clean_maint %>% select(year, usual_medicalcare_status, hispanic_status)
@@
 
-# Lets look at the smoking status among hispanic and non-hipanic women during the given years
+# Lets look at the smoking_status among hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_smoking_hisp <- data_clean_maint %>% select(year, smoking_status, hispanic_status)
@@
 
-# Lets look at the alcohol status among hispanic and non-hipanic women during the given years
+# Lets look at the alcohol_status hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_alcohol_hisp <- data_clean_maint %>% select(year, alcohol_status, hispanic_status)
@@
 
-# Lets look at the diabetes status among hispanic and non-hipanic women during the given years
+# Lets look at the diabetes_status hispanic_status (Hispanic and Non-hispanic women) during the given years
 year_diabetes_hisp <- data_clean_maint %>% select(year, diabetes_status, hispanic_status)
@@
 
-# Creating a demographic characteristics table on the basis of mammogram screening status
+# Creating a demographic characteristics table on the basis of mammogram_status
 table1 <- data_clean_maint %>% 
@@
 
-# Summarizing mammogram_status and Hispanic_status in the given years
+# Summarizing mammogram_status and hispanic_status in the given years
 figure1 <- ggplot(data_clean_main1, aes(x = mammogram_status)) +
@@
   scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000, 6000)) +
+  xlab ("Mammogram screening status of Hispanic and Non-hispanic women") +
+  ylab ("Count in numbers") +
+  ggtitle ("Mammogram status of study population in 2000, 2005, 2010, and 2015") +
   theme(legend.position = "bottom", 

### 2023-05-05 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Final Project Update

--- 2023-04-21 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
@@
 any(is.na(data_clean_main))
+
+# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
+data_clean_main$age <- as.factor(data_clean_main$age)
+data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
+data_clean_main$income <- as.factor(data_clean_main$income)
+data_clean_main$education <- as.factor(data_clean_main$education)
+data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
+data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
+data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
+data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
+data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
+data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
+data_clean_main$health_status <- as.factor(data_clean_main$health_status)
+data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
+data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
+data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
+data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
+data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
+
+# Recoding the variables
+data_clean_main$hispanic_status <- recode_factor(data_clean_main$hispanic_status, '1' = 'Non-Hispanic', '2' = 'Hispanic')
+data_clean_main$mammogram_status <- recode_factor(data_clean_main$mammogram_status, '0' = 'No', '1' = 'Yes')
+data_clean_main$age <- recode_factor(data_clean_main$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
+data_clean_main$income <- recode_factor(data_clean_main$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
+data_clean_main$education <- recode_factor(data_clean_main$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
+data_clean_main$marital_status <- recode_factor(data_clean_main$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
+data_clean_main$region_residence <- recode_factor(data_clean_main$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
+data_clean_main$insurance_status <- recode_factor(data_clean_main$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
+data_clean_main$health_status <- recode_factor(data_clean_main$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
+data_clean_main$usual_medicalcare_status <- recode_factor(data_clean_main$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
+data_clean_main$smoking_status <- recode_factor(data_clean_main$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_main$alcohol_status <- recode_factor(data_clean_main$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
+data_clean_main$diabetes_status <- recode_factor(data_clean_main$diabetes_status, '0' = 'No', '1' = 'Yes')
+
+# Limiting only 3 decimal places
+options(digits = 3)
 ```
@@
 log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
-log_fit_year
+tidy(log_fit_year)
 
 # Looking at the summary of log_fit_year
-tidy(log_fit_year)
+bivariate_year <- tidy(log_fit_year) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_year
+saveRDS(bivariate_year, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_year.rds")))
 ```
@@
 log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
-log_fit_hispanicstatus
+tidy(log_fit_hispanicstatus)
 
-# Looking at the summary of log_fit_hispanicstatus
-tidy(log_fit_hispanicstatus)
+# Looking and saving the summary of log_fit_hispanicstatus
+bivariate_hispanic <- tidy(log_fit_hispanicstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_hispanic
+saveRDS(bivariate_hispanic, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_hispanic.rds")))
 ```
@@
 log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
-log_fit_age
+tidy(log_fit_age)
 
@@
 tidy(log_fit_age)
+bivariate_age <- tidy(log_fit_age) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_age
+saveRDS(bivariate_age, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_age.rds")))
 ```
@@
 log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
-log_fit_income
+tidy(log_fit_income)
 
@@
 tidy(log_fit_income)
+bivariate_income <- tidy(log_fit_income) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_income
+saveRDS(bivariate_income, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_income.rds")))
 ```
@@
 log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)
-log_fit_education
 
@@
 tidy(log_fit_education)
+bivariate_education <- tidy(log_fit_education) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_education
+saveRDS(bivariate_education, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_education.rds")))
 ```
@@
 log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)
-log_fit_maritalstatus
 
@@
 tidy(log_fit_maritalstatus)
+bivariate_maritalstatus <- tidy(log_fit_maritalstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_maritalstatus
+saveRDS(bivariate_maritalstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_maritalstatus.rds")))
 ```
@@
 log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)
-log_fit_regionresidence
 
@@
 tidy(log_fit_regionresidence)
+bivariate_regionresidence <- tidy(log_fit_regionresidence) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_regionresidence
+saveRDS(bivariate_regionresidence, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_regionresidence.rds")))
 ```
@@
 log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)
-log_fit_insurancestatus
 
@@
 tidy(log_fit_insurancestatus)
+bivariate_insurancestatus <- tidy(log_fit_insurancestatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_insurancestatus
+saveRDS(bivariate_insurancestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_insurancestatus.rds")))
 ```
@@
 log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)
-log_fit_healthstatus
 
@@
 tidy(log_fit_healthstatus)
+bivariate_healthstatus <- tidy(log_fit_healthstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_healthstatus
+saveRDS(bivariate_healthstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_healthstatus.rds")))
 ```
@@
 log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)
-log_fit_usualmedicalcarestatus
 
@@
 tidy(log_fit_usualmedicalcarestatus)
+bivariate_usualmedicalcarestatus <- tidy(log_fit_usualmedicalcarestatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_usualmedicalcarestatus
+saveRDS(bivariate_usualmedicalcarestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_usualmedicalcarestatus.rds")))
 ```
@@
 log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)
-log_fit_smokingstatus
 
@@
 tidy(log_fit_smokingstatus)
+bivariate_smokingstatus <- tidy(log_fit_smokingstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_smokingstatus
+saveRDS(bivariate_smokingstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_smokingstatus.rds")))
 ```
@@
 log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)
-log_fit_alcoholstatus
 
@@
 tidy(log_fit_alcoholstatus)
+bivariate_alcoholstatus <- tidy(log_fit_alcoholstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_alcoholstatus
+saveRDS(bivariate_alcoholstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_alcoholstatus.rds")))
 ```
@@
 log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)
-log_fit_diabetesstatus
 
@@
 tidy(log_fit_diabetesstatus)
+bivariate_diabetesstatus <- tidy(log_fit_diabetesstatus) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+bivariate_diabetesstatus
+saveRDS(bivariate_diabetesstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_diabetesstatus.rds")))
 ```
@@
 log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
-log_fit_fullmodel
+tidy(log_fit_fullmodel)
 
-# Looking at the summary of log_fit_fullmodel
-table2 <- tidy(log_fit_fullmodel)
+# Looking and saving the summary of log_fit_fullmodel
+table2 <- tidy(log_fit_fullmodel) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
 table2
@@
 log_fit_year_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
-log_fit_year_age_hispanic
+tidy(log_fit_year_age_hispanic)
 
 # Looking at the summary of log_fit_year_age_hispanic
-table3 <- tidy(log_fit_year_age_hispanic)
+table3 <- tidy(log_fit_year_age_hispanic) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+table3
 saveRDS(table3, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds")))
@@
 log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
-log_fit_demographic
+tidy(log_fit_demographic)
 
 # Looking at the summary of log_fit_demographic
-table4 <- tidy(log_fit_demographic)
+table4 <- tidy(log_fit_demographic) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+table4
 saveRDS(table4, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds")))
@@
 log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
-log_fit_medical
+tidy(log_fit_medical)
 
 # Looking at the summary of log_fit_medical
-table5 <- tidy(log_fit_medical)
+table5 <- tidy(log_fit_medical) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+table5
 saveRDS(table5, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds")))
@@
 log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
-log_fit_smoking_alcohol
+tidy(log_fit_smoking_alcohol)
 
 # Looking at the summary of log_fit_smoking_alcohol
-table6 <- tidy(log_fit_smoking_alcohol)
+table6 <- tidy(log_fit_smoking_alcohol) %>% 
+  select(term, estimate, p.value) %>% 
+  mutate('odds ratio' = exp(estimate)) %>% 
+  select(term, 'odds ratio', p.value)
+table6
 saveRDS(table6, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds")))

### 2023-05-05 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
Final Project Update

@@
 roc_combined <- plot_grid(roc_fullmodel, roc_altmodel, ncol = 2, align = "h", axis = "tb",
-                           labels = c("ROC Curve for the full model", "ROC Curve for the alternative model"),
+                           labels = c("ROC Curve for full model", "ROC Curve for alternative model"),
                            label_size = 10)

### 2023-05-05 products/manuscript/Manuscript.qmd
Final Project Update

@@
 All bivariate and multivariate relationships between mammogram status and hypothesized predictors (year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status) were found to be significantly associated with mammogram screening.
-There was a higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
-Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users.
+There was a higher odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
+More educated women had higher odds of having mammogram compared to low or even non-educated women.
+Current smokers were found having lower odds of having mammograms compared to never smokers.
 
@@
 Multiple logistic regression was also performed for determining the multivariate effect on mammogram screening.
-Model selection process was performed and the final model model was selected based on ROC AUC (Area under the Receiver Operating Characteristic curve) value.
+Model selection process was performed and the final model was selected based on ROC AUC (Area under the Receiver Operating Characteristic curve) value.
 The model evaluation was performed using 5-fold cross-validation process repeated for 5 times.
@@
 Logistic regression was used for examining such relationships.
-In bivariate relationship (Supplementary Tables 3-12), we found statistically significant relationship between mammogram screening and hispanic status of women and found the log odds of having mammogram screening for Hispanic women as -0.56 compared to non-Hispanic women.
-The log odds of having mammogram was found as 0.25 in year 2015 compared to that of 2000.
-We also found that the log odds of having mammogram increased with age, and education.
-We found the log odds of having mammogram screening for women with income below the poverty threshold as -0.76 compared to women with income at or above the poverty threshold.
-The log odds of having mammogram for married was 0.67 compared to never married women.
-Women having the usual medical place had the log odds of having mammogram of 1.47 compared to those having no usual medical place.
+In bivariate relationship (Supplementary Tables 3-12), we found statistically significant relationship between mammogram screening and hispanic status of women and found the odds of having mammogram screening for Hispanic women as 0.57 compared to non-Hispanic women.
+The odds of having mammogram was found as 1.28 in year 2015 compared to that of 2000.
+We also found that the odds of having mammogram increased with age, and education.
+We found the odds of having mammogram screening for women with income below the poverty threshold as 0.47 compared to women with income at or above the poverty threshold.
+The odds of having mammogram for married women was 1.96 compared to never married women.
+Women having the usual medical place had the odds of having mammogram as 4.34 compared to those having no usual medical place.
 Similarly, Table 2 shows the multivariate relationship between mammogram screening and all other predictors of interests.
@@
 Table 3 shows the model performance comparison which compared the models' performance including various numbers of predictors.
-The best model performance was found for the full model including all predictors of interests (AIC: 16981 and RMSE: 0.32).
+The best model performance was found for the full model including all predictors of interests as it had the lowest AIC value of 16981 and RMSE value of 0.32.
 Similarly, based on model performance comparison, second best model was identified as the model including demographic predictors (age, hispanic status, income, education, marital status, and region of residence) that had AIC of 17764.1 and RMSE of 0.33.
@@
 ```
+*Note: The given model name includes mammogram status as a response variable and the following variables as predictors as given in brackets: log_fit_year (year), log_fit_hispanicstatus (hispanic status of women), log_fit_age (age), log_fit_income (income), log_fit_education (education), log_fit_maritalstatus (marital status), log_fit_regionresidence (region of residence), log_fit_insurancestatus (insurance status), log_fit_healthstatus (health status), log_fit_usualmedicalcarestatus (usual place for medical care), log_fit_smokingstatus (smoking status), log_fit_alcoholstatus (alcohol status), log_fit_diabetesstatus (diabetes status), log_fit_fullmodel (all predictors of interests), log_fit_year_age_hispanic (year, age, and hispanic status), log_fit_demographic (age, hispanic status, income, education, marital status, and region of residence), log_fit_medical (insurance status, health status, usual place for medical care, and diabetes status), and log_fit_smoking_alcohol (smoking status and alcohol status).
 
@@
 All bivariate and multivariate relationships between mammogram status and predictors were found to be significant hence, all hypothesized predictors were found to have strong association with mammogram screening.
-As suggested in previous studies (@katz_breast_2000), we found higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
-Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users.
+As suggested in previous studies (@katz_breast_2000), we found higher odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold.
+We also found older aged and highly educated women to have higher odds of having mammogram screening compared to young and low educated women.
+Current smokers were found having lower odds of having mammograms compared to people who never smoked.
 Previous studies have also suggested that people engaging in high risk behaviors are less likely to comply with cancer screening guidelines (@holm_health_1999).
@@
 This study confirms that there exists some association between mammogram screening and socio-economic status of Hispanic/non-Hispanic women during different years as all the hypothesized predictors were found to be significantly associated with mammogram status.
-We found age, education, income, and smoking/alcohol behavior as the great predictors for having mammogram screening.
+We found age, education, income, and smoking behavior as the great predictors for having mammogram screening.
 Further research needs to be done to dive more into these hypothesized predictors to see their absolute and relative effects on mammogram screening.

### 2023-05-05 products/supplement/Supplementary_Material.qmd
Final Project Update

@@
 ---
-title: "Supplement to Manuscript for MADA Project"
+title: "Supplementary Materials for Manuscript of MADA Project"
 subtitle: "Supplement Materials"
@@
 
-# Model Evaluation
-
-# Supplementary Table 20
-```{r, message = FALSE}
-#| echo: FALSE
- # Loading coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation
-result_table20 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
-result_table20
-```
-
-# Supplementary Table 21
-```{r, message = FALSE}
-#| echo: FALSE
- # Loading coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation
-result_table21 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
-result_table21
-```
-
-# Machine Learning (Model Selection)
-
-# Supplementary Table 22
-
-```{r, message = FALSE}
-#| echo: FALSE
- # Loading coefficients and penalty data during lasso model process
-result_table22 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
-result_table22
-```
-
 # List of Figures
